---
title: "Seurat - modified"
output:
  html_document:
    theme: united
    df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r rmd setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE
) 
```

```{r seurat object setup}
library(Seurat)
library(patchwork)
library(dplyr)
library(RColorBrewer)
library(DESeq2)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(tidyverse) 

# change work directory
CtxMg_23gw.data <- Read10X(data.dir ="/Users/jiapei/Desktop/single_cell_analysis/23gw/GeMg/filtered_feature_bc_matrix_GeMg_23gw/")
GeMg_23gw.data <- Read10X(data.dir ="/Users/jiapei/Desktop/single_cell_analysis/23gw/CtxMg/filtered_feature_bc_matrix_CtxMg_23gw/")

# Import scrublet (doublet detection package: https://github.com/swolock/scrublet), optional
CtxMg_23gw <- CreateSeuratObject(counts = CtxMg_23gw.data, project = "CtxMg_23gw", min.cells = 3, min.features = 200)
GeMg_23gw <- CreateSeuratObject(counts = GeMg_23gw.data, project = "GeMg_23gw", min.cells = 3, min.features = 200)

# Merge GE and CTX datasets
object <- merge(x = CtxMg_23gw, y = GeMg_23gw, merge.data = T, project = "object")
object
object[["percent.mt"]] <- PercentageFeatureSet(object = object, pattern = "^MT-")
object[["percent.rb"]] <- PercentageFeatureSet(object = object, pattern = "^RP[SL]")
head(x = object@meta.data, 5) # check metadata

wd <- "/Users/jiapei/Desktop/single_cell_analysis/github_pub" #change to the directory in which you want to save your data/graphs

setwd(wd)
# Save Seurat object locally
saveRDS(object, file = "object.rds")
```

```{r QC plots, Filter, anchor, pca}
setwd(wd)
# Plot QC
tiff("qcViolinPlot.tiff", width=1000, height=500, res=120)
VlnPlot(object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

tiff("qcFeaturePlot.tiff", width=1500, height=500, res=120)
plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") 
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt") 
plot3 <- FeatureScatter(object, feature1 = "percent.rb", feature2 = "percent.mt") 
CombinePlots(plots = list(plot1,plot2,plot3), ncol = 3)
dev.off()

# Based on QC data, filter Seurat object accordingly
object <- subset(x = object, subset = nFeature_RNA < 6000 & nCount_RNA < 30000 & percent.mt < 5)
object #check how many cells are left

# Additional filtering: only keep features (genes) that have >1 count in >5 cells, to save time in downstream analysis
mincounts<- 1
mincells <- 5
IsFeatureExpressed <- rowSums(as.matrix(object[["RNA"]]@counts >= mincounts)) >= mincells
ExpressedFeatures <- rownames(object)[IsFeatureExpressed]
object <- subset(x = object, features = ExpressedFeatures)
object

## The following steps use SCTransform 
## SCTransform combines NormalizeData, FindVariableFeatures, ScaleData. Results are saved in a new assay (can be custom named) with counts being (corrected) counts, data being log1p(counts), scale.data being pearson residuals; sctransform::vst intermediate results are saved in misc slot of this new assay.
# Step 1: Split object into a list of seurat objects
obj.list <- SplitObject(object, split.by = "orig.ident")
# Step 2: Normalize and identify variable features for each dataset independently
obj.list <- lapply(X = obj.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = obj.list, nfeatures = 2000)
obj.list <- PrepSCTIntegration(object.list = obj.list, anchor.features = features)
# Step 3: Select features that are repeatedly variable across datasets for integration; define integration anchors
anchors <- FindIntegrationAnchors(object.list = obj.list, normalization.method = "SCT", anchor.features = features)
combined.sct <- IntegrateData(anchorset = anchors, normalization.method = "SCT") 
# Step 4: Note that the original unmodified data still resides in the 'RNA' assay
DefaultAssay(combined.sct) <- "integrated"
# Step 5: Run the standard workflow for visualization and clustering
combined.sct <- RunPCA(combined.sct)

setwd(wd)
tiff("ElbowPlot.tiff", width=600, height=600, res=120)
ElbowPlot(object = combined.sct)
dev.off()

# Test different PCA # and resolution combos 
object <- combined.sct
DefaultAssay(object) <- "integrated"
object <- FindNeighbors(object, reduction = "pca", dims = 1:30) # try k.param ~1% of cell number
object <- FindClusters(object, resolution = 0.28)
head(x = Idents(object), 5)

object <- RunUMAP(object, reduction = "pca", dims = 1:30) #higher # of PCs form clearer cluster definitions
#combined.sct <- RunTSNE(combined.sct, reduction = "pca", dims = 1:15)
DimPlot(object, reduction = "umap")

# Save object!!
setwd(wd)
saveRDS(object, file = "Mg_ctrl_sample.rds")

# UMAP
setwd(wd)
tiff("umap_PC30_res028_1.tiff", width=2000, height=1000, res=120)
p1 <- DimPlot(object, reduction = "umap", group.by = "orig.ident") 
p2 <- DimPlot(object, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2
tiff("umap_PC30_res028_2.tiff", width=4000, height=1000, res=120)
DimPlot(object, reduction = "umap", split.by = "orig.ident")
dev.off()

# UMAP: QC
setwd(wd)
tiff("umap_nCount_legend.tiff", width=600, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "nCount_RNA") & NoAxes() #& NoLegend()
tiff("umap_nFeature_legend.tiff", width=600, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "nFeature_RNA") & NoAxes() #& NoLegend()
#tiff("umap_percentMt_legend.tiff", width=600, height=600, res=120)
#FeaturePlot(object, reduction = 'umap', features = "percent.mt") & NoAxes() #& NoLegend()
dev.off()

scrublet_num <- as.integer(as.logical(object@meta.data$scrublet))
object@meta.data$scrublet_num <- scrublet_num
setwd(wd)
tiff("umap_scrubletScore.tiff", width=600, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "scrublet_score") & NoAxes() & NoLegend()
tiff("umap_scrublet.tiff", width=600, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "scrublet_num") & NoAxes() & NoLegend()
dev.off()

# tSNE
setwd(wd)
tiff("tsne_origID_1_PC6_res02_kparam125.tiff", width=2000, height=1000, res=120)
p1 <- DimPlot(object, reduction = "tsne", group.by = "orig.ident")
p2 <- DimPlot(object, reduction = "tsne", label = TRUE, repel = TRUE)
p1 + p2
dev.off()

# PCA details
print(x = object[['pca']], dims = 1:5, nfeatures = 5)

setwd(wd)
tiff("PC_loading.tiff", width=1000, height=1000, res=120)
VizDimLoadings(object, dims = 1:2, reduction = 'pca')
#tiff("PCA_plot.tiff", width=750, height=500, res=120)
#DimPlot(object, reduction = 'pca')
#tiff("PC1to6_heatmap.tiff", width=1000, height=1000, res=120)
#DimHeatmap(object, dims = 1:6, cells = 500, balanced = T)
dev.off()

# calculate average reads and genes per cell
## method 1 ##
counts <- object@assays$RNA@counts # raw counts from Seurat object (after filtering)
counts_per_cell <- Matrix::colSums(counts)
print(mean(counts_per_cell)) # average reads per cell

counts_per_gene <- Matrix::rowSums(counts)
genes_per_cell <- Matrix::colSums(counts>0) # count gene only if it has non-zero reads mapped.
print(mean(genes_per_cell)) # average genes per cell

hist(log10(counts_per_cell+1),main='counts per cell',col='wheat')
hist(log10(genes_per_cell+1), main='genes per cell', col='wheat')
plot(counts_per_cell, genes_per_cell, log='xy', col='wheat')
title('counts vs genes per cell')

## method 2 ##
mean(object@meta.data$nCount_RNA) # average reads per cell
mean(object@meta.data$nFeature_RNA) # average genes per cell
```

```{r single dataset: SCTransform workflow}
object_endo <- SCTransform(object_endo, vars.to.regress = "percent.mt", variable.features.n = 2000)
object_endo <- RunPCA(object_endo) 

# if SCTransform takes too long, try this:
object <- NormalizeData(object = object, normalization.method = "LogNormalize", scale.factor = 1e4)
object <- FindVariableFeatures(object = object,selection.method = 'vst', nfeatures = 2000)
# Identify the 10 most highly variable genes
top10 <- head(x = VariableFeatures(object = object), 10)
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object = object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
all.genes <- rownames(x = object)
object <- ScaleData(object = object, features = all.genes)
# if scaling is taking too long, try this:
# default in `ScaleData` is only to perform scaling on the previously identified variable features (2,000 by default)
object <- ScaleData(object = object)
# to remove unwanted sources of variation from a single-cell dataset: we can 'regress out' heterogeneity associated with (for example) cell cycle stage, or mitochondrial contamination.
object <- ScaleData(object = object, vars.to.regress = 'percent.mt')
# perform linear dimentional reduction
object <- RunPCA(object = object, features = VariableFeatures(object = object))
```

```{r integrate data / batch correction 3: age and region plots}
object <- combined.sct
colnames(x = object[[]])
object@meta.data$age %>% table()

temp <- object@meta.data$age
temp[temp=="GW17"] <- "GW17_18"
temp[temp=="GW18"] <- "GW17_18"
temp[temp=="GW21"] <- "GW21_23"
temp[temp=="GW23"] <- "GW21_23"
object$age_group <- temp
object@meta.data$age_group %>% table()

#Endo <- subset(object, niche == "GeEndo" | niche == "CtxEndo")

setwd(wd) # to be plotted after confirming cluster identities
tiff("umap_CTX.tiff", width=400, height=400, res=120)
DimPlot(object_CTX, reduction = "umap") + NoAxes() + scale_color_manual(values=colors) + NoLegend()
dev.off()

# Save object!
setwd(wd)
saveRDS(object, file = "Mg_ctrl_sample.rds")
```

```{r integrate data / batch correction 4: find marker genes}
# For performing differential expression after integration, we switch back to the original data
DefaultAssay(object) <- "RNA"

markers <- FindAllMarkers(object, only.pos = T, min.pct = 0.5, logfc.threshold = 0.6) # find only positive markers
setwd(wd)
res <- markers %>% group_by(cluster) %>% top_n(n = 25, wt = avg_log2FC) #display top 25 genes in each cluster
write.csv(res, file = "top25marker.csv", row.names = T)
write.csv(markers, file = "markers.csv", row.names = T) # keep track of all DEGs

## Run additional DE analysis to differentiate similar clusters ##
test <- FindMarkers(object, ident.1 = "1 Homeostatic MG 1", ident.2 = "2 Homeostatic MG 2", min.pct = 0.5, logfc.threshold = 0.6)
setwd(wd)
write.csv(test, file = "1and2.csv", row.names = T)

## test ##
markers %>% group_by(cluster) %>% top_n(n = 12, wt = avg_log2FC) -> test
VlnPlot(object, features = subset(test,cluster == "7 APCs")$gene) 

# Plot marker genes in feature plot or violin plot
DefaultAssay(object) <- "SCT" # use SCT assay for plotting

top_selected <- c("P2RY12","AIF1","CCL2","GPNMB","MKI67","CLDN5","LYZ","HLA-DRA","STMN2","IL7R","MS4A1")

setwd(wd) 
tiff("FeatPlot_markers.tiff", width=1800, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = top_selected, 
            cols=c(adjustcolor("grey90", alpha.f = 0.2),"yellow","orange","red","darkred"), ncol = 6,
            keep.scale = "all") & NoAxes() & NoLegend() #pt.size = 0.5
#+ scale_colour_gradientn(colours = brewer.pal(n = 11, name = "YlOrRd"))
dev.off()

setwd(wd)
tiff("VlnPlot_CD14.tiff", width=2000, height=1000, res=120)
#Idents(object) <- "celltype"
VlnPlot(object, features = c("FCGR1A"), pt.size = 0) & NoLegend() #same.y.lims = T, cols = colors, y.max = 8, cols = c("black","red"), split.by = "condition", ncol = 3
dev.off()

# Plot markers genes in dot plot or ridge plot
setwd(wd)
tiff("Dotplot_markers.tiff", width=1200, height=600, res=120)
DotPlot(object, features = top_selected, dot.scale = 8, cols = c("blue", "red")) + RotatedAxis() # split.by = "age_label"
#tiff("RidgePlot_marker.tiff", width=1200, height=600, res=120)
#RidgePlot(object, feature = top_selected)
dev.off()

# Plot top marker genes as heatmap
markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) -> top5
top <- c("CX3CR1","P2RY12","AIF1","C1QA","VSIG4",
         "CCL2","CH25H","OTUD1","EGR2","NR4A2",
         "SPP1","GPNMB","TREM2","APOC1","APOE",
         "MKI67","TOP2A","CENPF","CDK1","AURKA",
         "CLDN5","IGFBP7","ID3","ITM2A","MFSD2A",
         "LYZ","S100A9","S100A12","JAML","ANXA1",
         "HLA-DRB1","HLA-DRA","HLA-DPA1","HLA-DPB1","CD74",
         "STMN2","TUBB2B","SOX4","NRXN1","DCX",
         "IL7R","GIMAP7","CD3D","TRBC1","TRAC",
         "MS4A1","IGLC1","IGHM","IGHD","IGLC2") 

# the following if based on Kwok et al, Immunity 2020 Fig 3
TF_monocyte <- c("IRF5","IRF8","KLF4","ZFYVE28","AHRR","GLI1","FOS","HOXA1","JUND","FOSL2","NFIL3","NRG1","JAG2","MEIS1","HOXA7",
                 "ZMAT3","SOX12","SOX4","MAML2","MLLT6","SCML2","PRDM5","PPARGC1B","TCF7L2","ETV5","TRPS1","POU2F2","PIR","RARG",
                 "FOSB","KLF2","JUN","IL31RA","GBX2","NOTCH4") #omitted: NOTCH4, GBX2, IL31RA, RARG, PIR, TRPS1, ETV5, PPARGC1B, PRDM5, SCML2, 
                  #MLLT6, MAML2, ZMAT3, HOXA7, MEIS1, JAG2, NRG1, HOXA1, GLI1, AHRR, ZFYVE28, IRF5
TF_neutrophil <- c("GFI1","CEBPE","PER3","JAG1","SOX13","KLF5","DACH1","ZEB1","ID2","ETS1","ELL2","HLF","ZBTB46","GFI1B","GATA1","KLF1",
                   "TCF7L1","POU2AF1","PTGER3","BATF3") #omitted: BATF3, PTGER3, POU2AF1, TCF7L1, KLF1, GATA1, GFI1B, ZBTB46, HLF, ELL2, DACH1, 
                  #KLF5, SOX13, CEBPE, GFI1
TF_GMP <- c("HOXA9","IKZF2","EYA2","HDX","ZFP36L1","GATA2","HOXA10","CIITA","ZFP57") #omitted: ZFP57, CIITA, HOXA10, GATA2, HDX, EYA2, IKZF2, HOXA9
# omission reason: not "varied" enough among cells/clusters
TF_combine <- c("IRF5","IRF8","KLF4","ZFYVE28","AHRR","GLI1","FOS","HOXA1","JUND","FOSL2","NFIL3","NRG1","JAG2","MEIS1","HOXA7",
                 "ZMAT3","SOX12","SOX4","MAML2","MLLT6","SCML2","PRDM5","PPARGC1B","TCF7L2","ETV5","TRPS1","POU2F2","PIR","RARG",
                 "FOSB","KLF2","JUN","IL31RA","GBX2","NOTCH4","GFI1","CEBPE","PER3","JAG1","SOX13","KLF5","DACH1","ZEB1","ID2",
                "ETS1","ELL2","HLF","ZBTB46","GFI1B","GATA1","KLF1","TCF7L1","POU2AF1","PTGER3","BATF3","HOXA9","IKZF2","EYA2","HDX","ZFP36L1",
                "GATA2","HOXA10","CIITA","ZFP57")
TF_combine_clean <- c("KLF4","FOSL2","NFIL3","ZFP36L1","TCF7L2","POU2F2","JUND","FOS","KLF2","FOSB","ID2","JUN",
                      "IRF8","SOX12","SOX4","PER3","JAG1","ZEB1","ETS1")

object_mo_neu <- subset(object, idents = c("5 Monocytes","9 Neutrophils"))
object_mo_neu_ctrl <- subset(object_mo_neu, subset = condition == "CTRL")
object <- subset(object_mo_neu, subset = condition == "GMH")

setwd(wd)
tiff("Heatmap_markers.tiff", width=1500, height=800, res=120)
DoHeatmap(object, features = top, label = F, group.colors = colors)# + NoLegend() #group.colors = c("brown","yellow")
dev.off()
```

```{r integrate data / batch correction 5: rename clusters}
IDs <- c("1b Homeostatic MG 2","1a Homeostatic MG 1","1c CCL2+ MG","2 WM-assoc MG","3 Cell cycle MG","4 Vascul-assoc MG","5 Monocytes","6 HLA+ My","7 Neuron-assoc MG","8 T cells","9 B cells") 
names(x = IDs) <- levels(x = object)
object <- RenameIdents(object, IDs)
object[["cell_type"]] <- Idents(object)
new_level <- c("1a Homeostatic MG 1","1b Homeostatic MG 2","1c CCL2+ MG","2 WM-assoc MG","3 Cell cycle MG","4 Vascul-assoc MG","5 Monocytes","6 HLA+ My","7 Neuron-assoc MG","8 T cells","9 B cells")
object@active.ident <- factor (x = object@active.ident, levels = new_level)

# Set up cluster colors (check google sheet)
colors <- c("blue","red","orange","forestgreen","cyan","grey50","darkorchid4","brown","magenta","yellow") #CCL2: bisque

# Plot clusters with new names
setwd(wd)
tiff("umap_PC30_res028.tiff", width=600, height=600, res=120)
DimPlot(object, reduction = "umap", label = F) + NoAxes() + NoLegend() + scale_color_manual(values=colors)
tiff("umap_PC30_res028_legend.tiff", width=600, height=600, res=120)
DimPlot(object, reduction = "umap") + NoAxes() + scale_color_manual(values=colors)
dev.off()

setwd(wd)
tiff("umap_age.tiff", width=2000, height=500, res=120)
DimPlot(object, reduction = "umap", split.by = "age") + NoAxes() + NoLegend() + scale_color_manual(values=colors)
dev.off()
```

```{r integrate data / batch correction 6: DEGs}
### DESeq2 ###
object_new <- subset(x = object, subset = celltype == "5 Vascul-assoc MG") 
Idents(object_new) <- "region" # parameter to run DESeq2 on
object_new@active.ident %>% table()
#object_subset <- subset(object, orig.ident != "9 Cytotoxic T and NK cells")
#object_DEG <- object_subset
object_new[["RNA"]]@counts <- as.matrix(object_new[["RNA"]]@counts)+1 #only use to debug error in following line
DEG <- FindMarkers(object_new, ident.1 = "GE", ident.2 = "CTX", assay = "RNA", test.use = "DESeq2", max.cells.per.ident = 1000)
# max.cells.per.ident default is Inf (no downsampling)
# try test.use = "MAST" --> MAST produces a weird volcano plot

setwd("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/previous/by_sample")
write.csv(DEG, file = "GEvsCTX_VAM_DESeq2.csv", row.names = T)

# order csv file by logFC; volcano plot:
setwd(wd)
DE_genes <- read.csv("GMHvsCTRL_DESeq2_mo.csv")
#min_adj <- 0
min_adj <- min(DE_genes$p_val_adj[DE_genes$p_val_adj > 0], na.rm = T); min_adj
#DE_genes[is.na(DE_genes)] <- 0 # replace any NA values with 0
for (row in 1:dim(DE_genes)[1])
  {
    DE_genes$p_val_adj[row] <- -log10(DE_genes$p_val_adj[row] + min_adj) #prevent log transform 0
  }
n <- 5
genes.to.label.2 <- c(DE_genes$X[1:n], tail(DE_genes$X,n)) #extract first and last n DE genes (ranked by log2FC)
genes.to.label <- c("CD14","RNASE1","C1QB","S100A9","GAPDH","CXCL16")
label.index <- match(genes.to.label,DE_genes$X)
#genes.to.label.2 <- c("CX3CR1","P2RY12","CCL2","OLR1","CH25H")
label.index.2 <- match(genes.to.label.2,DE_genes$X) [6:10]
  
df <- data.frame(x = DE_genes$avg_log2FC, y = DE_genes$p_val_adj)
f <- ggplot(df, aes(x,y)) + 
  geom_point(colour = "black", size = 1) +
  #geom_point(aes(colour = cut(DE_genes$p_val_adj, c(-Inf, 1.3, Inf)))) + # -log(0.05) = 1.30103
  #scale_color_manual(name = "FDR cutoff (<0.05)", values = c("(-Inf,1.3]" = "black", "(1.3, Inf]" = "red"),
  #labels = c(">= 0.05", "< 0.05")) +
  scale_y_continuous(name = "-log10 Adjusted p-value", limits = c(0,300)) +
  scale_x_continuous(name = "log2 Fold Change", limits = c(-4,4)) + 
  geom_point(data = df[label.index,], aes(x,y), colour = "red", size = 2) +
  geom_label_repel(data = df[label.index,], aes(x,y,label=DE_genes$X[label.index]), size = 3, color = "red") +
  geom_point(data = df[label.index.2,], aes(x,y), colour = "blue", size = 2) +
  geom_label_repel(data = df[label.index.2,], aes(x,y,label=DE_genes$X[label.index.2]), size = 3, color = "blue") +
  theme(axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
#f <- f + geom_hline(yintercept=1.3, linetype="dashed", color = "red")
#f <- f + geom_label_repel(aes(label=ifelse(All_genes$log2FoldChange < -9, as.character(All_genes$symbol),'')),size = 3, color = "red")
#f <- f + geom_label_repel(aes(label=ifelse(All_genes$log2FoldChange > 29, as.character(All_genes$symbol),'')),size = 3, color = "red")
f

setwd(wd)
tiff("DESeq2_volcano_GMHvsCTRL_mo.tiff", width=700, height=700, res=120)
f
dev.off()

library(EnhancedVolcano)
EnhancedVolcano(DEGs, lab = rownames(DEGs), x = "avg_log2FC", y = "p_val", pCutoff = 1e-09, FCcutoff = 0.5, labSize = 5, 
                legendPosition = "none", drawConnectors = F, title = "title")

### avg count plot ###
theme_set(theme_cowplot())

object <- SetIdent(object, value = 'celltype')
object_subset <- subset(object, celltype == "5 Vascul-assoc MG")
avg_expr <- log1p(AverageExpression(object = object_subset)$RNA) # change accordingly
avg_expr <- as.data.frame(avg_expr) #switch matrix into data frame
avg_expr$gene <- rownames(avg_expr)
avg_expr <- na.omit(avg_expr) #eliminate NaN rows

avg_expr_1 <- avg_expr[avg_expr$CTX > 10,] #select high expr genes (to eliminate technical errors)
avg_expr_1 <- avg_expr_1[avg_expr_1$CTX - avg_expr_1$GE > 30,] #select genes with higher expr in this subset
avg_expr_2 <- avg_expr[avg_expr$GE > 10,]
avg_expr_2 <- avg_expr_2[avg_expr_2$GE - avg_expr_2$CTX > 30,]

genes.to.label.1 <- c("CCL4L2","CCL4","CCL3L1","SPP1","CCL3")#avg_expr_1$gene
genes.to.label.2 <- c("HBG1","HBB","TMSB10") #avg_expr_2$gene
label.index.1 <- match(genes.to.label.1,avg_expr$gene)
label.index.2 <- match(genes.to.label.2,avg_expr$gene)
label.index.avg.1 <- match(avg_expr_1$gene,avg_expr$gene)
label.index.avg.2 <- match(avg_expr_2$gene,avg_expr$gene)

p1 <- ggplot(avg_expr, aes(CTX, GE)) + geom_point() + 
  geom_point(data = avg_expr[label.index.avg.1,], aes(CTX, GE), colour = "red", size = 2) +
  geom_label_repel(data = avg_expr[label.index.1,], aes(CTX, GE, label=avg_expr$gene[label.index.1]), size = 3, color = "red") +
  geom_point(data = avg_expr[label.index.avg.2,], aes(CTX, GE), colour = "red", size = 2) +
  geom_label_repel(data = avg_expr[label.index.2,], aes(CTX, GE, label=avg_expr$gene[label.index.2]), size = 3, color = "red") +
  scale_y_continuous(limits = c(0,700)) +
  scale_x_continuous(limits = c(0,700))
#p1 <- LabelPoints(plot = p1, points = genes.to.label, repel = T)
p1

setwd(wd)
write.csv(avg_expr, file = "avg_expr_region_5.csv", row.names = T)  ## change accordingly

setwd(wd)
tiff("avgExpr_region_5.tiff", width=500, height=500, res=120)
p1
dev.off()

# for loop to process DE genes in each cell type
setwd(wd) 
#levels(x = object)
mg <- c(0,1,3,4,8,11)
endo <- c(2,5,7,13,14,15)
for (i in endo)
{
  test.markers <- FindMarkers(object, ident.1 = i, ident.2 = endo[!endo %in% i], min.pct = 0.25)
  write.csv(test.markers, file = paste("MgEndo_sp_", i, ".csv", sep=""), row.names = T)
}

# Make violin plots in a loop
gene_list <- list(E14, P4, E14_P4, peripheral, DAM, lr_candidates,lr_candidates_selected)
names(gene_list) <- c("E14","P4","E14_P4","peripheral","DAM","lr_candidates","lr_candidates_selected")

i <- 1
for (list in gene_list){
  setwd(wd)
  tiff(paste("VlnPlot_",names(gene_list)[i],".tiff", sep=""), width=3000, height=ceiling(length(list)/3)*600, res=120)
  print(VlnPlot(object, features = list))
  dev.off()
  i <- i+1
}

# find DE genes with roc (not very accurate)
setwd(wd)
Mg_23gw.markers_roc <- FindAllMarkers(object = object, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
res_roc <- Mg_23gw.markers_roc %>% group_by(cluster) %>% top_n(n = 10, wt = myAUC) #display top 10 genes in each cluster
write.csv(res_roc, file = "Mg_23gw_top10marker_roc.csv", row.names = T) #rank by ROC scores; not as accurate

# Save object!
setwd(wd)
saveRDS(object, file = "scvelo_mg_sample.rds")
```

```{r integrate data / batch correction 7: VS other datasets}
query <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/external_data/ChrisGlass_mg_12gw/Mg_12gw.rds")
query_endo <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/18_23_endo/endo.rds")
DefaultAssay(query_endo) <- "RNA"
reference <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/scvelo_mg_sample.rds")
DefaultAssay(reference) <- "RNA"

reference <- RunUMAP(reference, reduction = "pca", dims = 1:15, return.model = T)

test <- ProjectUMAP(query = reference, query.reduction = "pca", reference = reference, reference.reduction = "pca", reduction.model = "umap")

DimPlot(test, reduction = "ref.umap", group.by = "celltype")
```

```{r integrate data / batch correction 8: VS bulk seq}
install.packages("readxl")
library("readxl")
HLA <- c("TRAF6","CD74","HLA-DPB1","HLA-DMB","HLA-DMA","LGMN","B2M","HLA-DQA2","HLA-DRB1","HLA-DQB1","CTSV","PIKFYVE","CTSE",
         "F6UB75","HLA-DRB5","FCGR2B","HLA-DOA","CTSF","CTSD","HLA-DRB4","HLA-DQB2","HLA-DPA1","FCER1G","HLA-DOB","HLA-DRB4",
         "HLA-DQA1","HLA-DRA","IFI30","DNM2","CTSL","CTSS","HLA-DRB3")
edgeR_region <- read_excel("/Users/jiapei/Desktop/UCSF_BMS/bulkRNAseq/edgeR/results/ctxVSge/region_ctxVSge.xlsx")
edgeR_GE_up_logFC <- subset(edgeR_region,logFC>0)
edgeR_GE_up_logFC <- edgeR_GE_up_logFC$Symbol
#edgeR_GE_up_filtered <- subset(edgeR_GE_up_logFC,FDR<0.05)$Symbol
edgeR_GE_up_HLA <- intersect(edgeR_GE_up_logFC,HLA)

edgeR_GE_up <- c("SOCS1","APCDD1","AZU1","CLIP4","FN1","CLEC14A","LGALS3BP","SOX18")
DESeq2_GE_up <- c("MORN3","CLEC4C","HHLA1","CCDC81","S1PR3","SOX18","GZMA","ZNF16","PCP4","MSX1","CLEC14A","ZNF311","KIRREL1","PTGIR",
                  "TIMP4","PRSS3","PRICKLE1","PCDHGC3","C20orf204","ADH1B","SSMEM1","IGIP","KLKB1","TMEM225","SLC6A9","RARRES2","PP2D1",
                  "ZNF324B","CLEC12B","FAM228A","ATP1B4","KCNJ8","NCR3LG1","RGS7BP","C2CD4C","ZIC1","ACTL9","C8orf37","ZNF132","P2RY4",
                  "CYP7A1","SUV39H1","IQCF2","IYD","NIM1K","RAMP3","PKDCC","DZIP1L","LGALS3BP","CYP4F11","TBXA2R","C22orf34","CPA3",
                  "ETAA1","ELN","SUSD5","EHD1","GPR61","ASPHD2","EFNA4","CCN2","HORMAD1","CASTOR1","SOCS1","SPON1","NOB1","FOXS1","BEND5",
                  "MSLN","ZNF484","TEF","GMPPA","BYSL","GUCY1A2","HFE","RCE1","LAMC1","CDA","MYC","ADAMTSL3","EGR4","SEMA4B","OCA2","DUSP4",
                  "GIPC3","LSR","DGKB","GAREM2","STARD3","ZDHHC11B","LMF1","NKILA","MCAM","CRHBP","COL4A2","ZFP90","CPA6","CD34",
                  "ZNF366","ZNF304","PLCL1","TLCD2","NATD1","EXD3","JAG2","KDR","DCUN1D5","NOTCH3","AMPD2","RASGEF1C","ATIC","SLC9B1",
                  "CC2D2B","SLC27A5","FZR1","MFSD2A","MCC","DBP","CRY2","GDPGP1","SPINT1","IKZF4","COL5A2","PLEKHG1","ZNF468","PDE7B",
                  "GORAB","HSD17B8","PPM1F","ZNF767P","PPP1R16A","ADGRL3","RBM42","LAMA5","CXXC5","PLEKHA8","BZW2","ABAT","SLC16A10",
                  "DRG2","FGFR1","ACTN4","B3GAT3","PNMA1","GLTP","THBS3","ECE1","ETS1","MATK","GIMAP6","MTURN","DDR1","TEX14","FN1",
                  "C12orf65","UBE4B","PAPSS2","SEPTIN9","SLC7A8","DCLRE1C","DROSHA","NFKBIE")
edgeR_CTX_up <- c("APOBEC3B","PNLDC1","PCDH11Y","ZNF202","OASL","C16orf70","ORC5","WDR24","NEIL2","TCEANC","CPXM1","CDCA5","RHBDD1",
                  "ARMC8","SPP1","GEN1","ZNF34","SLC25A13","FBXO16","TMEM138","VDAC3","NOM1","SIMC1","PCP4L1","FZD8","AZI2","SEC22B")
DESeq2_CTX_up <- c("TRIM37","YEATS2","IMPDH2","GNPTAB","MXD1","FANCI","CPSF1","DCLK1","SOCS3","MAP1B","CRMP1","RASA2","ZC3H8","GEN1",
                   "TCTN1","MPLKIP","GALNT6","NGDN","SUN1","ANAPC15","Corf89","BDH1","EFHC1","KHNYN","CAPRIN2","AP2A1","NCOA5","CXADR",
                   "STEAP1B","DCUN1D","PRAM1","FBXO22","MTMR10","DPAGT1","MAP2K4","RBMS2","ACOT9","AAAS","PGRMC1","PPP1R8","RRM1",
                   "TBL2","ACAD8","ZNF426","PBK","KNL1","TTC21A","POC1A","POMGNT2","RPP40","ZSCAN21","SULT4A1",
                   "RWDD2B","MAPK4","PRR7","ASS1","TBX15","CETP","RMDN2","IGSF9","C7orf25","PCDHGA7","CTRB2")
edgeR_1519_up <- c("CDC20","GMCL1","NIM1K","METTL7B","ACTR3B","VCX","TARS2","CCNG1","C3orf38","PTEN","ZNF567","TSC22D3","IQCB1","NCAPH",
                   "ODF2","RMI2","SLC66A3","AZU1","INMT","CAPN7","GGT1","OLA1","VPS13B","PREB","DLEU7","C11orf1","HMBS","LYAR","DLD",
                   "ECI1","TRIM31","ANGPTL1","SCCPDH","VWA5A","SNPH","PCDH11Y","PNLDC1","FBXL17","STX18","NUP93","GPR101","PAAF1",
                   "FAM49B","HOXC8","COPS4","ALPK3","ETFDH","DERA","GSDME","GUSBP5","TIGD4","UBA3","MT1A","BRINP2","PDRG1","HMGN4",
                   "ARMC10","C14orf177","ODF3","GMPR2","C1orf56","GID4","PDE2A","SMAGP","ZNF202","GEN1","RNASEH2B","MORF4L2","DHRS1",
                   "PSMC2","ARHGAP29","SIMC1","SLC30A5","FERMT2","MKLN1","INTS10","SMC1A","SGO1","ABHD12","MUC1","NUDT16","SEC23B",
                   "FGD3","CLTB","OR5K4","PCP4L1","ORC5","OASL","B4GALT3","CCDC90B","GCH1","TEX10","TPMT","ERG28","PIGF","CCT8ARSB",
                   "C6orf47","NEFL","GRPEL1","TEX14","VPS25","KMT5C","FZD8","SIDT1","C8orf48","CCIN","ARMC8","DCUN1D5","TMEM138","WDR24",
                   "DEFA4","H3C13","STX12","REXO1","ARMC1","TRPV2","ANKRD40CL","CD68","SERPINB7","CCDC190","TRIP10","RNF14","NDUFB6",
                   "AKIRIN1","SPG21","PAQR8","PLSCR1","VCPKMT","SIGLEC14","HCG9","GLRX5","MGST2","CHEK1","TXNDC9","TCEANC","OR6K6","VEZF1",
                   "RGS2","DIS3","MAMDC4","KLHDC8B","TNFRSF13C","PAIP2B","RRP7BP","SLX4IP","ELOVL3","SURF2","WASHC3","SEC22A","PAG1","PCMT1",
                   "FUCA2","NEIL2","ARHGAP18","TIMM50","FBLN1","CDKN3","ATP6V1H","POLR2H","TRIM52","CRY2","CCDC130","TMSB4X","FBXO16","CPXM1",
                   "FBXO2","APOBEC3B","BPIFB3","BPIFB1","ZNF449","SH3KBP1","RBBP4","LRIF1","TSPAN5","MDH1","UFC1","CAHM","RHBDD1","SDC2",
                   "RGS1","MCC","ST6GAL2","SEC22B","ATOH7")
edgeR_2123_up <- c("HAS1","IGFBP2","CCN2","STRA6","NEFL","APLNR","GLRB","CCN1","BCHE","TMEM204","CTHRC1","FN1","C4orf36","SOX18","GADD45A")

common <- intersect(edgeR_GE_up_filtered,rownames(object)) # change accordingly
#all(common %in% rownames(object)) # check that all genes are in the Seurat object
object[["GE_up_bulk_HLA_edgeR"]] <- PercentageFeatureSet(object, features = common)
object[["GE_up_bulk_edgeR_nofilter"]] <- PercentageFeatureSet(object, features = common)
object[["GE_up_bulk_edgeR_filter"]] <- PercentageFeatureSet(object, features = common)
object[["GE_up_bulk_edgeR"]] <- PercentageFeatureSet(object, features = common)
object[["CTX_up_bulk_edgeR"]] <- PercentageFeatureSet(object, features = common)
object[["X1519_up_bulk_edgeR"]] <- PercentageFeatureSet(object, features = common)
object[["X2123_up_bulk_edgeR"]] <- PercentageFeatureSet(object, features = common)

setwd(wd)
tiff("UMAP_GE_up_bulk_filter_edgeR.tiff", width=700, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "GE_up_bulk_edgeR_filter", 
            cols=c(adjustcolor("grey90", alpha.f = 0.2),"red","darkred","black")) & NoAxes() #& NoLegend() 
tiff("VlnPlot_GE_up_bulk_filter_edgeR.tiff", width=600, height=600, res=120)
VlnPlot(object,features = "GE_up_bulk_HLA_edgeR", cols = colors, pt.size = 0) & NoLegend() 
dev.off()

library(ggplot2)
library(scales)
library(ggnewscale)

umapCoord <- as.data.frame(Embeddings(object[["umap"]]))
GE_up <- object@meta.data$GE_up_bulk_HLA_edgeR
GE_up_log <- log(GE_up + min(GE_up[GE_up > 0]))
CTX_up <- object@meta.data$CTX_up_bulk_edgeR
CTX_up_log <- log(CTX_up + min(CTX_up[CTX_up > 0]))
GE_up_log %>% summary()
CTX_up_log %>% summary()
# Combine with original data and blended colours
df_test <- cbind.data.frame(
  setNames(umapCoord, c("x", "y")), GE_up, CTX_up,
  colour = rgb(
    rescale(GE_up_log),
    0,
    rescale(CTX_up_log)))

for (i in 1:nrow(df_test)) {
  if (df_test[i,]$GE_up < 0.8 & df_test[i,]$CTX_up < 12) {
    df_test[i,]$colour <- rgb(211/255,211/255,211/255,0.05) # setting cells below certain threholds to transparent grey
  }
}

setwd(wd)
tiff("UMAP_region_bulk_edgeR.tiff", width=800, height=600, res=120)
ggplot(df_test, aes(x, y, colour = colour)) +
  geom_point() +
  scale_colour_identity() +
  new_scale_colour() + # shape = NA --> invisible layers
  geom_point(aes(colour = GE_up_log), shape = NA) +
  scale_colour_gradient(low = "black", high = "red") +
  new_scale_colour() +
  geom_point(aes(colour = CTX_up_log), shape = NA) +
  scale_colour_gradient(low = "black", high = "blue") +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),
        panel.background=element_blank(),panel.border=element_blank(),
        panel.grid.major=element_blank(),panel.grid.minor=element_blank(),plot.background=element_blank()) # keep graph background blank
dev.off()

# AddModuleScore (ref: https://github.com/satijalab/seurat/issues/3521)
# does not perform as well as PercentageFeatureSet (see aboce)
DefaultAssay(object) <- 'RNA'
edgeR_GE_up_list <- list(edgeR_GE_up)
object <- AddModuleScore(object = object, features = edgeR_2123_up_list, name = "edgeR_2123_up_ModScore")

setwd(wd)
tiff("UMAP_2123_up_bulk_edgeR_ModScore.tiff", width=700, height=600, res=120)
FeaturePlot(object, reduction = 'umap', features = "edgeR_2123_up_ModScore1", 
            cols=c(adjustcolor("grey90", alpha.f = 0.2),"red","darkred","black")) & NoAxes()
dev.off()

setwd(wd)
saveRDS(object, file = "scvelo_mg_sample.rds")
```

```{r integrate data / batch correction 9: filter DEG files}
DEG <- read_excel("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/GEvsCTX_DESeq2.xlsx")
DEG$p_val_adj <- as.numeric(DEG$p_val_adj) # in case column is recognized as character
DEG_filter <- subset(DEG, p_val_adj < 0.05)

library("xlsx")
setwd(wd)
write.xlsx(DEG_filter, "GEvsCTX_DESeq2_filter.xlsx", col.names = T, row.names = T)
```

```{r cell type composition plot}
# cell type composition for each donor
object <- combined.sct
#object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/17gw_mg_anchor/17_mg_anchor_SCT.rds")

# Create a stacked percent barplot (by cell type)
data <- data.frame(matrix(ncol=3, nrow=1, dimnames=list(NULL, c("Cell_type", "Freq", "Condition"))))
for (cell_type_ID in levels(object@meta.data$celltype)) 
{
  meta_subset <- subset(object@meta.data, celltype == cell_type_ID) 
  freq <- as.data.frame(table(meta_subset$orig.ident)) ## change accordingly
  freq$perc <- freq$Freq / sum(freq$Freq)
  for (var in freq$Var1)
  {
    data <- rbind(data, c(cell_type_ID, subset(freq,Var1==var)$perc, var))
  }
}
data <- data[2:dim(data)[1],] #eliminate the first row (dummy row to make rbind work)
data$Freq <- as.numeric(data$Freq)
# make different levels for each cell subtype
level_order <- data[['Cell_type']]
level_order <- level_order[!duplicated(level_order)]

setwd(wd)
tiff("barplot_cellType_age.tiff", width=1000, height=600, res=120)
ggplot(data, aes(fill = Condition, y = Freq, x = factor(Cell_type, level = level_order))) + 
  geom_bar(position="stack", stat="identity") +
  xlab("Cell subtypes") +
  ylab("Age frequency") +
  scale_x_discrete(breaks = level_order, labels = level_order) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_discrete(name = "Sample", labels = c("17GW CTX","17GW GE"))
dev.off()

# Create a stacked percent barplot (by sample)
#object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/scvelo/by_sample/scvelo_mg_sample.rds")
#object <- subset(x = object, subset = sample == "GW23_2")
data <- data.frame(matrix(ncol=3, nrow=1, dimnames=list(NULL, c("Cell_type", "Freq", "Condition"))))
for (parameter in c("CTX","GE")) ## change accordingly
{
  meta_subset <- subset(object@meta.data, region == parameter) ## change accordingly
  freq <- as.data.frame(table(meta_subset$cell_type)) ## change parameter accordingly (e.g. to celltype)
  freq$perc <- freq$Freq / sum(freq$Freq)
  for (var in freq$Var1)
  {
    data <- rbind(data, c(var, subset(freq,Var1==var)$perc*100, parameter))
  }
}
data <- data[2:dim(data)[1],] #eliminate the first row (dummy row to make rbind work)
data$Freq <- as.numeric(data$Freq)

#data$Condition2 <- data$Condition
#data$Condition2[data$Condition2=="GeMg_23gw"] <- "CTRL" ## change accordingly
#data$Condition2[data$Condition2=="GeMg_GMH_24gw"] <- "GMH" ## change accordingly
level_order <- data[['Condition']]
level_order <- level_order[!duplicated(level_order)]

setwd(wd)
write.csv(data, file = "barplot_region.csv", row.names = T)  ## change accordingly

setwd(wd)
tiff("barplot_HLA_region.tiff", width=500, height=650, res=120) ## change accordingly
ggplot(data, aes(fill = factor(Cell_type), y = Freq, x = factor(Condition))) + # factor(Cell_type, levels=new_level)
  geom_bar(position="fill", stat="identity") +
  xlab("Condition") +
  ylab("% cells in cluster") +
  scale_x_discrete(breaks = level_order, labels = level_order) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10), axis.text.y = element_text(size = 10),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  guides(fill=guide_legend(title="Cluster")) +
  scale_y_continuous(labels = scales::percent_format()) 
  #scale_fill_manual(values=colors)
dev.off()

setwd(wd)
tiff("barplot_cluster_condition.tiff", width=1000, height=600, res=120) ## change accordingly
ggplot(data, aes(fill = factor(Condition), y = Freq, x = factor(Cell_type, levels=new_level))) +
  geom_bar(position="dodge", stat="identity") +
  xlab("Cell type") +
  ylab("% cells in cluster") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10), axis.text.y = element_text(size = 10),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  guides(fill=guide_legend(title="Condition"))
dev.off()

## make scatterplot of cluster %
setwd(wd)
data <- read.csv(file = "barplot_region_scatterplot.csv")

setwd(wd)
tiff("scatterplot_cluster_region.tiff", width=400, height=400, res=120) ## change accordingly
ggplot(data, aes(x=CTX, y=GE)) + geom_point() + geom_label(label=rownames(data)) + geom_abline(colour = "blue") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
dev.off()
```

```{r additional plots}
# New things to try!  Note that plotting functions now return ggplot2 objects, so you can add themes, titles, and options
FeaturePlot(object = object, features = c("AIF1", "CLDN5"), blend = TRUE)
DimPlot(object = object) + DarkTheme()
DimPlot(object = object) + labs(title = "23GW microglia and endothelial cells")

# Plot with ggplot2
plot_umap <- DimPlot(object = object, reduction = "umap", label = F, label.size = 4.5) + xlab("UMAP 1") + ylab("UMAP 2") + 
  theme(axis.title = element_text(size = 18), legend.text = element_text(size = 18)) + 
  guides(colour = guide_legend(override.aes = list(size = 10)))
ggsave(filename = "/Users/jiapei/Desktop/single_cell_analysis/23gw/Mg_23gw_umap_ggplot.png", height = 7, width = 12, plot = plot_umap)

plot_tsne <- DimPlot(object = object, reduction = "tsne", label = F, label.size = 4.5) + xlab("tSNE 1") + ylab("tSNE 2") + 
  theme(axis.title = element_text(size = 18), legend.text = element_text(size = 18)) + 
  guides(colour = guide_legend(override.aes = list(size = 10)))
ggsave(filename = "/Users/jiapei/Desktop/single_cell_analysis/23gw/Mg_23gw_tsne_ggplot.png", height = 7, width = 12, plot = plot_tsne)
```

```{r export data}
# Save raw counts (need to convert gene names to ensembl ID)
#BiocManager::install("EnsDb.Hsapiens.v79")
library(EnsDb.Hsapiens.v79)

count_mtx <- as.matrix(GetAssayData(object = object, slot = "counts"))
dim(count_mtx)
count_mtx[1,1:5] 

geneSymbols <-  rownames(count_mtx)
rownames(count_mtx)[1:5]
ensembl_ID <- ensembldb::select(EnsDb.Hsapiens.v79, keys= geneSymbols, keytype = "SYMBOL", columns = c("SYMBOL","GENEID"))
#not all genes have matching ensembl IDs
count_mtx_ensembl <- count_mtx[ensembl_ID$SYMBOL,]
dim(count_mtx_ensembl)[1]==dim(ensembl_ID)[1] #should have same number of rows as length of ensembl_ID; return T
rownames(count_mtx_ensembl)[1:5]==ensembl_ID$SYMBOL[1:5] #should return T
count_mtx_ensembl <- cbind.data.frame(Gene = ensembl_ID$GENEID,count_mtx_ensembl) #add ensembl ID to the first column
dim(count_mtx_ensembl) #should have 1 additional column than prev
head(count_mtx_ensembl,3) #first 3 genes (matrix now turned into data frame)

write.table(count_mtx_ensembl,'/Users/jiapei/Desktop/single_cell_analysis/18gw/Mg_18gw_counts_ensembl.txt', 
            sep = '\t', row.names = F, col.names = T, quote = F)

# Save raw counts (gene names)
# as.matrix(GetAssayData(object = object, slot = "data"))[200:210,10:12] == as.matrix(GetAssayData(object = object, slot = "counts")) [200:210,10:12]
endo <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/CTRL_endo/ctrl_endo.rds")
mg <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/scvelo_mg_sample.rds")
object <- merge(x = endo, y = mg, merge.data = T, project = "MgEndo")
object #check celltypes are what you expect

setwd(wd)
saveRDS(object,"MgEndo.rds")

setwd(wd)
count_raw <- as.matrix(GetAssayData(object_2023_GE, slot = "counts"))
write.table(count_raw, 'raw_counts_2023_GE.txt', 
            sep = '\t', row.names = T, col.names = T, quote = F) #raw count data (not transformed)

setwd(wd)
count_norm <- apply(count_raw, 2, function(x) (x/sum(x))*10000)
write.table(count_norm, 'norm_counts.txt', 
            sep = '\t', row.names = T, col.names = T, quote = F) #normalized data (rec by CellPhoneDB)

setwd(wd)
write.table(as.matrix(GetAssayData(object_2023_GE, slot = "data")), 'norm_counts_seurat_2023_GE.txt', 
            sep = '\t', row.names = T, col.names = T, quote = F) #normalized data by Seurat

# if you want to downsize number of cells in count mtx (subset by cell type first)
object@meta.data$cluster %>% table()
mono2 <- subset(x = object, subset = cluster == "mono2")
count_raw_mono2 <- as.matrix(GetAssayData(mono2, slot = "counts"))
count_norm_mono2 <- apply(count_raw_mono2, 2, function(x) (x/sum(x))*10000)

subset <- sample(1:dim(count_norm_cd4)[2], 2000, replace = F)
count_norm_cd4_subset <- count_norm_cd4[,subset]
dim(count_norm_cd4_subset)

setwd(wd)
write.table(count_norm_cd4_subset, 'norm_counts_cd4.txt', sep = '\t', row.names = T, col.names = T, quote = F)

# Save metadata table (with new Cluster IDs)
colnames(x = object[[]])
object[["celltype"]] <- Idents(object = object)
head(x = object@meta.data, 5) #check whether new cell type names are added

setwd(wd)
write.table(object_2023_GE@meta.data, 'metadata_2023_GE.txt', sep = '\t', row.names = T, col.names = NA, quote = F)
# downsize version
#write.table(cd4@meta.data[subset,], 'metadata_cd4.txt', sep = '\t', row.names = T, col.names = NA, quote = F)

# Save gene names
gene_List <- rownames(object)
setwd(wd)
write.table(gene_List, 'features.tsv', sep = '\t', row.names = T, col.names = NA, quote = F)
```

```{r object interaction}
#ref: https://satijalab.org/seurat/articles/essential_commands.html 
library(tidyverse)

## Explore ##

colnames(x = object) %>% head() #cell names
rownames(x = object) %>% head() #feature (gene) names
ncol(x = object) #number of cells
nrow(x = object) #number of genes
# Get cell identity classes
levels(x = object)
Idents(object = object) %>% head()
# Set identity classes to an existing column in meta data
Idents(object = object) <- "orig.ident"
# Rename identity classes
object <- RenameIdents(object = object, `Endo-U` = "Endo-U1")
levels(x = object)

## Subset ##

# Subset Seurat object based on identity class, also see ?SubsetData
subset(x = object, idents = "Mg-U1")
subset(x = object, idents = c("Mg-U1", "Mg-Prolif"), invert = TRUE)
# Subset on the expression level of genes/features
subset(x = object, subset = MS4A1 > 3)
subset(x = object, subset = MS4A1 > 3 & PC1 > 5)
subset(x = object, subset = MS4A1 > 3, idents = "B cells")
# Subset on a value in the object meta data
subset(x = object, subset = orig.ident == "Replicate1")
# Downsample the number of cells per identity class
subset(x = object, downsample = 100)

## Access data ##

# View metadata data frame, stored in object@meta.data
object[[]] %>% head()
# Retrieve specific values from the metadata
object$nCount_RNA %>% head()
object[[c("percent.mt", "nFeature_RNA")]] %>% head()
# Add metadata, see ?AddMetaData
region_label <- sample(x = c("g1", "g2"), size = ncol(x = pbmc), replace = TRUE)
object$groups <- region_label
# Retrieve or set data in an expression matrix ('counts', 'data', and 'scale.data')
GetAssayData(object = object, slot = "counts")
object <- SetAssayData(object = object, slot = "scale.data", new.data = new.data)
# Get cell embeddings and feature loadings
Embeddings(object = object, reduction = "pca") %>% head()
Loadings(object = object, reduction = "pca") %>% head()
#Loadings(object = object, reduction = "pca", projected = TRUE)
# FetchData can pull anything from expression matrices, cell embeddings, or metadata
FetchData(object = object, vars = c("PC_1", "percent.mt", "MS4A1")) %>% head()

## Visualize ##

# Scatter plot
FeatureScatter(object = object, feature1 = "AIF1", feature2 = "PC_1")
FeatureScatter(object = object, feature1 = "AIF1", feature2 = "CLDN5")
FeatureScatter(object = object, feature1 = "AIF1", feature2 = "SPP1")
# HoverLocator show extra data throught the `information` argument, designed to work smoothly with FetchData
plot <- DimPlot(object = object) + NoLegend()
HoverLocator(plot = plot, information = FetchData(object = object, vars = c("ident", "PC_1", "nFeature_RNA")))
# FeatureLocator
select.cells <- CellSelector(plot = plot)
# Label points on a ggplot object
LabelPoints(plot = plot, points = TopCells(object = object[["pca"]]), repel = TRUE)
```

```{r published data: Hammond 2019}
library(dplyr)
library(Seurat)
library(readr)

counts.B10 <- read.delim2("/Users/jiapei/Downloads/GSM3442006_E14_F_B10.dge.txt", row.names=1)
object.B10 <- CreateSeuratObject(counts = counts.B10, project = "B10_F", min.cells = 5, min.features = 200)

counts.B11 <- read.delim2("/Users/jiapei/Downloads/GSM3442007_E14_M_B11.dge.txt", row.names=1)
object.B11 <- CreateSeuratObject(counts = counts.B11, project = "B11_M", min.cells = 5, min.features = 200)

counts.B12 <- read.delim2("/Users/jiapei/Downloads/GSM3442008_E14_F_B12.dge.txt", row.names=1)
object.B12 <- CreateSeuratObject(counts = counts.B12, project = "B12_F", min.cells = 5, min.features = 200)

counts.C1 <- read.delim2("/Users/jiapei/Downloads/GSM3442009_E14_F_C1.dge.txt", row.names=1)
object.C1 <- CreateSeuratObject(counts = counts.C1, project = "C1_F", min.cells = 5, min.features = 200)

counts.B9 <- read.delim2("/Users/jiapei/Downloads/GSM3442010_E14_M_B9.dge.txt", row.names=1)
object.B9 <- CreateSeuratObject(counts = counts.B9, project = "B9_M", min.cells = 5, min.features = 200)

counts.B6 <- read.delim2("/Users/jiapei/Downloads/GSM3442011_E14_F_B6.dge.txt", row.names=1)
object.B6 <- CreateSeuratObject(counts = counts.B6, project = "B6_F", min.cells = 5, min.features = 200)

counts.B7 <- read.delim2("/Users/jiapei/Downloads/GSM3442012_E14_M_B7.dge.txt", row.names=1)
object.B7 <- CreateSeuratObject(counts = counts.B7, project = "B7_M", min.cells = 5, min.features = 200)

counts.B8 <- read.delim2("/Users/jiapei/Downloads/GSM3442013_E14_M_B8.dge.txt", row.names=1)
object.B8 <- CreateSeuratObject(counts = counts.B8, project = "B8_M", min.cells = 5, min.features = 200)

object.B10
object.B11
object.B12
object.C1
object.B9
object.B6
object.B7
object.B8

combine_Mg <- merge(x = object.B10, y = c(object.B11,object.B12,object.C1,object.B9,object.B6,object.B7,object.B8), 
                    merge.data = T, project = "combine_Mg")
combine_Mg
object <- combine_Mg #assign object accordingly!
object[["percent.mt"]] <- PercentageFeatureSet(object = object, pattern = "^mt.")
head(x = object@meta.data, 5)

wd <- "/Users/jiapei/Desktop/single_cell_analysis/hammond/"

setwd(wd)
tiff("qc.tiff", width=2000, height=1000, res=120)
VlnPlot(object = object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
dev.off()
```

```{r cellID}
# ref: https://github.com/RausellLab/CelliD 
# https://bioconductor.org/packages/devel/bioc/vignettes/CelliD/inst/doc/BioconductorVignette.html 
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("ggpubr")) install.packages("ggpubr")
BiocManager::install("CelliD")

library(CelliD)
library(tidyverse) # general purpose library for data handling
library(ggpubr) #library for plotting

# Input: RDS object (from Seurat)
object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/17_18_23_anchor_new/Mg_SCT.rds") #already normalized, filtered, batch-corr

# Restricting to protein-coding genes
data("HgProteinCodingGenes")
object <- object[rownames(object) %in% HgProteinCodingGenes,]

# Normalize and scale
#object <- NormalizeData(object)
#object <- ScaleData(object, features = rownames(object))

# Multiple Correspondence Analysis (MCA)
object <- RunMCA(object)

wd <- "/Users/jiapei/Desktop/single_cell_analysis/17_18_23_anchor_new"
setwd(wd)
tiff("cellID_MCAplot.tiff", width=1200, height=1000, res=120)
DimPlotMC(object, reduction = "mca")
#DimPlotMC(object, reduction = "mca", group.by = "orig.ident", features = c("TMEM119", "CCL2"), as.text = T)
dev.off()

# Comparison: PCA
object <- RunPCA(object, features = rownames(object))
object <- RunUMAP(object, dims = 1:15)
object <- RunTSNE(object, dims = 1:15)
PCA  <- DimPlot(object, reduction = "pca")  + ggtitle("PCA") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
tSNE <- DimPlot(object, reduction = "tsne")+ ggtitle("tSNE") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
UMAP <- DimPlot(object, reduction = "umap") + ggtitle("UMAP") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
MCA <- DimPlot(object, reduction = "mca")  + ggtitle("MCA") + theme(legend.text = element_text(size =10), aspect.ratio = 1)

setwd(wd)
tiff("cellID_MCA_PCA.tiff", width=2000, height=1000, res=120)
ggarrange(PCA, MCA, common.legend = T, legend = "top")
tiff("cellID_tSNE_UMAP.tiff", width=2000, height=1000, res=120)
ggarrange(tSNE, UMAP, common.legend = T, legend = "top")
dev.off()

# Cell type prediction using pre-established marker lists
panglao <- read_tsv("https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz")

# restricting the analysis to brain specific gene signatues
panglao_blood <- panglao %>% filter(organ == "Blood")
# restricting to human specific genes
panglao_blood <- panglao_blood %>%  filter(str_detect(species,"Hs"))
# converting dataframes into a list of vectors, which is the format needed as input for CelliD
panglao_blood <- panglao_blood %>%  
  group_by(`cell type`) %>%  
  summarise(geneset = list(`official gene symbol`))
blood_gs <- setNames(panglao_blood$geneset, panglao_blood$`cell type`)

# restricting the analysis to brain specific gene signatues
panglao_brain <- panglao %>% filter(organ == "Brain")
# restricting to human specific genes
panglao_brain <- panglao_brain %>%  filter(str_detect(species,"Hs"))
# converting dataframes into a list of vectors, which is the format needed as input for CelliD
panglao_brain <- panglao_brain %>%  
  group_by(`cell type`) %>%  
  summarise(geneset = list(`official gene symbol`))
brain_gs <- setNames(panglao_brain$geneset, panglao_brain$`cell type`)

#filter to get human specific genes
panglao_all <- panglao %>%  filter(str_detect(species,"Hs"))
# convert dataframes to a list of named vectors which is the format for CelliD input
panglao_all <- panglao_all %>%  
  group_by(`cell type`) %>%  
  summarise(geneset = list(`official gene symbol`))
all_gs <- setNames(panglao_all$geneset, panglao_all$`cell type`)
#remove very short signatures
all_gs <- all_gs[sapply(all_gs, length) >= 10]

# Performing per-cell hypergeometric tests against the gene signature collection
HGT_blood_gs <- RunCellHGT(object, pathways = blood_gs, dims = 1:50, n.features = 200)
HGT_brain_gs <- RunCellHGT(object, pathways = brain_gs, dims = 1:50, n.features = 200)

# For each cell, assess the signature with the lowest corrected p-value (max -log10 corrected p-value)
blood_gs_prediction <- rownames(HGT_blood_gs)[apply(HGT_blood_gs, 2, which.max)]
# For each cell, evaluate if the lowest p-value is significant
blood_gs_prediction_signif <- ifelse(apply(HGT_blood_gs, 2, max)>2, yes = blood_gs_prediction, "unassigned")
# Save cell type predictions as metadata within the Seurat object
object$blood_gs_prediction <- blood_gs_prediction_signif

# For each cell, assess the signature with the lowest corrected p-value (max -log10 corrected p-value)
brain_gs_prediction <- rownames(HGT_brain_gs)[apply(HGT_brain_gs, 2, which.max)]
# For each cell, evaluate if the lowest p-value is significant
brain_gs_prediction_signif <- ifelse(apply(HGT_brain_gs, 2, max)>2, yes = brain_gs_prediction, "unassigned")
# Save cell type predictions as metadata within the Seurat object
object$brain_gs_prediction <- brain_gs_prediction_signif

# Comparing the original labels with CelliD cell-type predictions based on pancreas-specific gene signatures
OriginalPlot <- DimPlot(object, reduction = "umap", group.by = "celltype") + scale_color_manual(values=colors) +
  theme(legend.text = element_text(size = 8), aspect.ratio = 1)

setwd(wd)
tiff("celltype_predict_panglao_blood.tiff", width=2000, height=1000, res=120)
Predplot1 <- DimPlot(object, reduction = "umap", group.by = "blood_gs_prediction") + 
  theme(legend.text = element_text(size = 10), aspect.ratio = 1)
ggarrange(OriginalPlot, Predplot1, legend = "top")
tiff("celltype_predict_panglao_brain.tiff", width=2000, height=1000, res=120)
Predplot2 <- DimPlot(object, reduction = "umap", group.by = "brain_gs_prediction") + 
  theme(legend.text = element_text(size = 10), aspect.ratio = 1)
ggarrange(OriginalPlot, Predplot2, legend = "top")
dev.off()

# To evaluate all cell types
HGT_all_gs <- RunCellHGT(object, pathways = all_gs, dims = 1:50)
all_gs_prediction <- rownames(HGT_all_gs)[apply(HGT_all_gs, 2, which.max)]
all_gs_prediction_signif <- ifelse(apply(HGT_all_gs, 2, max)>2, yes = all_gs_prediction, "unassigned")
object$all_gs_prediction <- ifelse(all_gs_prediction_signif %in% c(names(blood_gs), "Platelets", "Erythoroid-like and erythroid precursor cells", "unassigned"), all_gs_prediction_signif,"other")

DimPlot(object, group.by = "all_gs_prediction", reduction = "umap") + theme(legend.text = element_text(size = 10), aspect.ratio = 1)

# Extracting per-cell gene signatures from my own dataset with CelliD(c)
cell_gs <- GetCellGeneSet(object, dims = 1:50, n.features = 200)
# Extracting per-group gene signatures from my own dataset with CelliD(g)
group_gs <- GetGroupGeneSet(object, dims = 1:50, n.features = 200, group.by = "celltype")

library('biomaRt')

eggen44 <- temp_obj #start with i in 1 and comment out merge line
for (i in 2:12) ##
{
  mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
  temp <- read.delim2(paste("/Users/jiapei/Desktop/single_cell_analysis/eggen/GSE141862_RAW/44_",i,".txt", sep="")) ##
  gene <- temp$gene
  
  G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol"), values=gene, mart= mart)
  temp <- merge(temp, G_list, by.x="gene", by.y="ensembl_gene_id")
  temp <- temp[!duplicated(temp$hgnc_symbol),]
  
  dim2 <- dim(temp)[2]
  temp2 <- temp[,-dim2]
  temp2 <- temp2[,-1]
  rownames(temp2) <- temp[,dim2]
  temp2 <- temp2[!row.names(temp2)%in%c(""),] #for some reason, one of the row names is empty (throwing error in CreateSeuratObject)
  
  temp_obj <- CreateSeuratObject(counts = temp2, project = paste("eggen44_",i,sep=""), min.cells = 5) ##
  eggen44 <- merge(x = eggen44, y = temp_obj, merge.data = T, project = "eggen44") ##
}

VlnPlot(eggen44, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2) ##
eggen44@meta.data$orig.ident <- "eggen44" ##

# merge all 20 samples and SCT!
eggen <- merge(x = eggen05, y = c(eggen09,eggen18,eggen19,eggen20,eggen22,eggen23,eggen25,eggen26,eggen28,eggen29,eggen30,eggen31,eggen32,eggen34,eggen36,eggen37,eggen41,eggen42,eggen44), merge.data = T, project = "eggen")

obj.list <- SplitObject(eggen, split.by = "orig.ident")
obj.list <- lapply(X = obj.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = obj.list, nfeatures = 2000)
obj.list <- PrepSCTIntegration(object.list = obj.list, anchor.features = features)
anchors <- FindIntegrationAnchors(object.list = obj.list, normalization.method = "SCT", anchor.features = features)
combined.eggen <- IntegrateData(anchorset = anchors, normalization.method = "SCT") 

# specify that downstream analysis on the corrected data; note that the original unmodified data still resides in the 'RNA' assay
DefaultAssay(combined.eggen) <- "integrated"

# Run the standard workflow for visualization and clustering
combined.eggen <- RunPCA(combined.eggen)

setwd(wd)
tiff("ElbowPlot.tiff", width=600, height=600, res=120)
ElbowPlot(object = combined.eggen)
dev.off()

# Test different PCA # and resolution combos
combined.eggen <- FindNeighbors(object = combined.eggen, reduction = "pca", dims = 1:30) 
combined.eggen <- FindClusters(object = combined.eggen, resolution = 0.5)

head(x = Idents(object = combined.eggen), 5)

combined.eggen <- RunUMAP(combined.eggen, reduction = "pca", dims = 1:30) #forms clearer cluster definitions
combined.eggen <- RunTSNE(combined.eggen, reduction = "pca", dims = 1:30)

# Save object!!
setwd(wd)
saveRDS(combined.eggen, file = "combined_eggen.rds")

# Visualization 1
setwd(wd)
tiff("umap_origID_1_PC30_res05.tiff", width=2000, height=1000, res=120)
p1 <- DimPlot(combined.eggen, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(combined.eggen, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2
tiff("tsne_origID_1_PC30_res05.tiff", width=2000, height=1000, res=120)
p1 <- DimPlot(combined.eggen, reduction = "tsne", group.by = "orig.ident")
p2 <- DimPlot(combined.eggen, reduction = "tsne", label = TRUE, repel = TRUE)
p1 + p2
dev.off()

data("HgProteinCodingGenes")
combined.eggen <- combined.eggen[rownames(combined.eggen) %in% HgProteinCodingGenes,]

combined.eggen <- RunMCA(combined.eggen, nmcs = 50)

combined.eggen <- RunPCA(combined.eggen)
combined.eggen <- RunUMAP(combined.eggen, dims = 1:30)
combined.eggen <- RunTSNE(combined.eggen, dims = 1:30)
PCA  <- DimPlot(combined.eggen, reduction = "pca")  + ggtitle("PCA") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
tSNE <- DimPlot(combined.eggen, reduction = "tsne")+ ggtitle("tSNE") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
UMAP <- DimPlot(combined.eggen, reduction = "umap") + ggtitle("UMAP") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
MCA <- DimPlot(combined.eggen, reduction = "mca")  + ggtitle("MCA") + theme(legend.text = element_text(size =10), aspect.ratio = 1)

setwd(wd)
tiff("cellID_MCA_PCA.tiff", width=2000, height=1000, res=120)
ggarrange(PCA, MCA, common.legend = T, legend = "top")
tiff("cellID_tSNE_UMAP.tiff", width=2000, height=1000, res=120)
ggarrange(tSNE, UMAP, common.legend = T, legend = "top")
dev.off()

# cell to cell matching
HGT_cell_gs <- RunCellHGT(combined.eggen, pathways = cell_gs, dims = 1:50)

cell_gs_match <- rownames(HGT_cell_gs)[apply(HGT_cell_gs, 2, which.max)]
cell_gs_prediction <- object$celltype[cell_gs_match]
cell_gs_prediction_signif <- ifelse(apply(HGT_cell_gs, 2, max)>2, yes = cell_gs_prediction, "unassigned")
combined.eggen$cell_gs_prediction <- cell_gs_prediction_signif

setwd(wd)
tiff("cell_cell_predict.tiff", width=1000, height=1000, res=120)
DimPlot(combined.eggen, group.by = "cell_gs_prediction", pt.size = 0.2, reduction = "umap") +
  ggtitle("Predicitons") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
dev.off()

# group to cell matching
HGT_group_gs <- RunCellHGT(combined.eggen, pathways = group_gs, dims = 1:50)

group_gs_prediction <- rownames(HGT_group_gs)[apply(HGT_group_gs, 2, which.max)]
group_gs_prediction_signif <- ifelse(apply(HGT_group_gs, 2, max)>2, yes = group_gs_prediction, "unassigned")
combined.eggen$group_gs_prediction <- group_gs_prediction_signif

setwd(wd)
tiff("group_cell_predict.tiff", width=1000, height=1000, res=120)
DimPlot(combined.eggen, group.by = "group_gs_prediction", pt.size = 0.2, reduction = "umap") + 
  ggtitle("Predicitons") + theme(legend.text = element_text(size =10), aspect.ratio = 1)
dev.off()

# KEGG pathway
# Downloading functional gene sets: For computational reasons, we just developed here the assessment on Hallmark MSigDB gene sets that is provided in the package. Other functionnal pathways can be obtained using the msigdbr package (didn't run this block yet)
library(msigdbr)
msigdf <- msigdbr::msigdbr(species = "Homo sapiens")
msigdf_filtered <- msigdf %>% dplyr::select(gs_cat, gs_subcat, gs_name, gene_symbol, gs_description) %>%
  filter(gs_subcat == "CP:KEGG") %>%
  mutate(gs_subcat = ifelse(gs_subcat == "", "Hallmark", gs_subcat))
KEGGgenesetDF <- msigdf_filtered %>%  mutate(gs_name = str_replace(sub("^.*?_{1}", "", gs_name), "_", "\\-")) %>%
  group_by(gs_subcat, gs_name) %>%
  summarise(geneset = list(gene_symbol)) %>%
  group_by(gs_subcat) %>%
  summarise(all_geneset = list(setNames(geneset, gs_name))) 
KEGGgeneset <- setNames(KEGGgenesetDF$all_geneset, KEGGgenesetDF$gs_subcat)

# Assessing per-cell functional enrichment analyses
HGT_Hallmark <- RunCellHGT(object, pathways = Hallmark, dims = 1:50)
HGT_KEGG <- RunCellHGT(object, pathways = KEGGgeneset$`CP:KEGG`, dims = 1:50)

object@assays[["Hallmark"]] <- CreateAssayObject(HGT_Hallmark)
object@assays[["KEGG"]] <- CreateAssayObject(HGT_KEGG)
# Visualizing per-cell functional enrichment annotations in  MCA, PCA, tSNE, UMAP
g1 <- FeaturePlot(object, "WNT-SIGNALING-PATHWAY", order = T, reduction = "umap", cols = c("grey", "red")) + 
      theme(legend.text = element_text(size =10), aspect.ratio = 1)
g1

setwd(wd)
tiff("FeatPlot_VEGF.tiff", width=500, height=500, res=120)
g1
dev.off()

```

```{r pseudobulk}
#ref: https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html 
library(SingleCellExperiment)
library(scater)
library(Seurat)
library(tidyverse)
library(cowplot)
library(Matrix.utils)
library(edgeR)
library(dplyr)
library(magrittr)
library(Matrix)
library(purrr)
library(reshape2)
library(S4Vectors)
library(tibble)
library(SingleCellExperiment)
library(pheatmap)
library(apeglm)
library(png)
library(DESeq2)
library(RColorBrewer)
library(xlsx)
library(readxl)

seurat <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/Final/CTRL_mg_171823gw/Mg_17_18_23gw.rds")

# Extract raw counts and metadata to create SingleCellExperiment object
counts <- seurat@assays$RNA@counts 
colnames(seurat@meta.data)
metadata <- seurat@meta.data 

# Set up metadata as desired for aggregation and DE analysis
#metadata$cluster_id <- factor(seurat@meta.data$celltype)
metadata$sample_id <- factor(seurat@meta.data$orig.ident)
# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), colData = metadata)
# Identify groups for aggregation of counts
groups <- colData(sce)[, c("sample_id")]

# Explore the raw counts for the dataset
## Explore the raw counts for the dataset
dim(counts(sce))
counts(sce)[1:6, 1:6]
## Explore the cellular metadata for the dataset
dim(colData(sce))
head(colData(sce))

# Named vector of sample names
sids <- purrr::set_names(levels(sce$sample_id))
# Total number of samples 
ns <- length(sids); ns

# Generate sample level metadata
## Determine the number of cells per sample
table(sce$sample_id)
## Turn named vector into a numeric vector of number of cells per sample
n_cells <- as.numeric(table(sce$sample_id)); n_cells
## Determine how to reorder the samples (rows) of the metadata to match the order of sample names in sids vector
m <- match(sids, sce$sample_id)
## Create the sample level metadata by combining the reordered metadata with the number of cells corresponding to each sample.
ei <- data.frame(colData(sce)[m, ], n_cells, row.names = NULL); ei

# Aggregate the counts per sample_id and cluster_id
# Subset metadata to only include the cluster and sample IDs to aggregate across
groups <- colData(sce)[, c("sample_id")] ## change to only sample_id 
# Aggregate across cluster-sample groups
pb <- aggregate.Matrix(t(counts(sce)), groupings = groups, fun = "sum") 
class(pb); dim(pb)
pb[, 1:10]

# Generate count mtx and sample table (based on metadata)
cluster_counts <- t(pb); cluster_counts[1:10,]
cluster_metadata <- read_excel("/Users/jiapei/Desktop/single_cell_analysis/pseudobulk/sample_table.xlsx")
rownames(cluster_metadata) <- cluster_metadata$orig.ident
# Check that all row names of metadata are the same and in the same order as column names of counts in order to use as input to DESeq2
all(rownames(cluster_metadata) == colnames(cluster_counts)) # should return TRUE

# Generate DGE matrix for DESeq2
dds <- DESeqDataSetFromMatrix(cluster_counts, 
                              colData = cluster_metadata, 
                              design = ~ region + age)

# Transform counts for data visualization
rld <- rlog(dds, blind=T)
# Plot PCA
wd <- "/Users/jiapei/Desktop/single_cell_analysis/pseudobulk"
setwd(wd)
tiff("DESeq2_PCA_mg_region.tiff", width=500, height=500, res=120)
DESeq2::plotPCA(rld, intgroup = "region")
dev.off()

# Hierarchical clustering
# Extract the rlog matrix from the object and compute pairwise correlation values
rld_mat <- assay(rld)
rld_cor <- cor(rld_mat)
# Plot heatmap
setwd(wd)
tiff("DESeq2_heatmap_mg.tiff", width=700, height=500, res=120)
pheatmap(rld_cor, annotation = cluster_metadata[, c("region_label"), drop=F])
dev.off()

# Run DESeq2 differential expression analysis
dds <- DESeq(dds)
# Plot dispersion estimates
plotDispEsts(dds)
```

```{r monocle3 setup}
# NEED TO RUN FROM MONOCLE3 ENVIRONMENT
if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")
BiocManager::install(version = "3.10") # NEED TO RUN FROM R3 ENVIRONMENT

BiocManager::install(c('BiocGenerics', 'DelayedArray', 'DelayedMatrixStats',
                       'limma', 'S4Vectors', 'SingleCellExperiment',
                       'SummarizedExperiment', 'batchelor', 'Matrix.utils'))

#install.packages("devtools", dependencies=T) # does not work in R 3.6.1
library(devtools) # ended up installing devtools using conda install -c r r-devtools 
# more: https://stackoverflow.com/questions/20923209/problems-installing-the-devtools-package 
devtools::install_github('cole-trapnell-lab/leidenbase') #having lots of trouble installing this on mac (try linux at lab?)
devtools::install_github('cole-trapnell-lab/monocle3')

remotes::install_github('satijalab/seurat-wrappers')

devtools::install_github('satijalab/seurat-data')

library(monocle3)
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(magrittr)
```

```{r scvelo}
library(Seurat)
library(SeuratDisk)
library(SeuratWrappers)
library(devtools)
library(tidyverse)
#install_github("velocyto-team/velocyto.R")

ldat <- ReadVelocity(file = "/Users/jiapei/Desktop/single_cell_analysis/scvelo/run_count_new_ctrl_GeMG.loom")
GeMg_21gw <- as.Seurat(x = ldat)
GeMg_21gw[["RNA"]] <- GeMg_21gw[["spliced"]]

# change orig.ident
object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/new_20221024/by_origID/Mg_ctrl.rds")
object@meta.data$orig.ident %>% table()

GeMg_21gw_10x <- subset(object, subset = orig.ident == "GeMg_21gw")
colnames(GeMg_21gw_10x) %>% head(3)
colnames(GeMg_21gw) %>% head(3)

cell_filter <- colnames(x = GeMg_21gw_10x)
cell_filter <- str_replace(cell_filter, "-1_2", "x")
cell_filter <- paste0("run_count_new_ctrl_GeMG:", cell_filter)
# filter out cells with >5% mitochondrial genes; append percent.mt
GeMg_21gw <- subset(GeMg_21gw, cells = cell_filter)
GeMg_21gw@meta.data$percent.mt <- GeMg_21gw_10x@meta.data$percent.mt
GeMg_21gw@meta.data$orig.ident <- "GeMg_21gw"
GeMg_21gw@meta.data %>% head(2)
nrow(GeMg_21gw@meta.data) == nrow(GeMg_21gw_10x@meta.data) #should return True

GeMg_GMH_25gw <- subset(object, subset = orig.ident == "GeMg_GMH_25gw")
GeMg_GMH_25gw_scvelo <- subset(object_scvelo, subset = orig.ident == "GeMg_GMH_25gw")
#CtxMg_23gw@meta.data %>% head(3)
#CtxMg_23gw_scvelo@meta.data %>% head(3)
# filter out cells with >5% mitochondrial genes; append percent.mt & scrublet scores (not sure if this step is necessary)
GeMg_GMH_25gw_scvelo@meta.data$percent.mt <- GeMg_GMH_25gw@meta.data$percent.mt # added here because for some reason percent.mt is 0 after loom (so appended percent.mt from before loom conversion)
GeMg_GMH_25gw_scvelo@meta.data$scrublet <- GeMg_GMH_25gw@meta.data$scrublet
GeMg_GMH_25gw_scvelo@meta.data$scrublet_num <- GeMg_GMH_25gw@meta.data$scrublet_num
GeMg_GMH_25gw_scvelo@meta.data %>% head(3)

# run merge pipeline (see above) and save as separate RDS object?
setwd("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/new_20221024")
DefaultAssay(object) <- "RNA"
SaveH5Seurat(object, filename = "Mg_ctrl_20221024.h5Seurat")
Convert("Mg_ctrl_20221024.h5Seurat", dest = "h5ad") # then go to python!
```

```{r cellphoneDB dotplot}
library(tidyverse)
library(dplyr)
library(ggplot2)
# input files
wd = "/Users/jiapei/Desktop/single_cell_analysis/GMHvsCTRL_mgendo/MgEndo_origID_subsample3000_threshold25"
setwd(wd)
pvalues <- read.delim("pvalues.txt", header = T, sep = "\t")
means <- read.delim("means.txt", header = T, sep = "\t")

# define subgroups of interacting cell types
interacting_cells <- 12:ncol(pvalues)
CTX <- grep("CtxMg.*.CtxEndo|CtxEndo.*.CtxMg", colnames(pvalues)) 
GE <- grep("GeMg.*.GeEndo|GeEndo.*.GeMg", colnames(pvalues))

MgHomeo <- grep("Homeostatic.MG.*.ndo|Homeostatic.MG.*.Tip|Tip.*.Homeostatic.MG|ndo.*.Homeostatic.MG", colnames(pvalues))
VAM <- grep("Vascul.assoc.MG.*.ndo|Vascul.assoc.MG.*.Tip|Tip.*.Vascul.assoc.MG|ndo.*.Vascul.assoc.MG", colnames(pvalues))
HLA <- grep("HLA.*.ndo|HLA.*.Tip|Tip.*.HLA|ndo.*.HLA", colnames(pvalues))
T_cell <- grep("T.cells.*.ndo|T.cells.*.Tip|Tip.*.T.cells|ndo.*.T.cells", colnames(pvalues))
Cytotoxic <- grep("Cytotoxic.*.ndo|Cytotoxic.*.Tip|Tip.*.Cytotoxic|ndo.*.Cytotoxic", colnames(pvalues))
Neutrophils <- grep("Neutrophils.*.ndo|Neutrophils.*.Tip|Tip.*.Neutrophils|ndo.*.Neutrophils", colnames(pvalues))

interaction <- GE #change accordingly

# pre-defined ligand-receptor pairs
#setwd(wd)
#selected_lr <- read.delim("rows_selected_EndoMg.txt", header = F, sep = "\t")
#selected_rows <- match(t(selected_lr),pvalues$interacting_pair)
# no selection on ligand-receptor pairs
selected_rows <- 1:dim(pvalues)[1]

selected_pvalues <- pvalues[selected_rows,interaction]
selected_pvalues <- -log(selected_pvalues+0.001,10) #-log10 transform; added 0.001 to avoid INF
filter <- rowSums(selected_pvalues, na.rm = T)>1 #exclude ligand-receptor pairs with no sig (pval = 1)
selected_pvalues <- selected_pvalues[filter,] 

selected_means <- means[selected_rows,interaction] #expression level (normalized)
selected_means <- log(selected_means+0.9993071,2) #alog2 transform; added constant to avoid INF (min: -0.001)
selected_means <- selected_means[filter,]
#dim(selected_pvalues) == dim(selected_means)

selected_rows <- selected_rows[filter]

selected_df <- data.frame()
for (i in 1:length(selected_rows)){
  temp_df <- data.frame(cell_type = colnames(pvalues)[interaction], gene_pair = pvalues[selected_rows[i],]$interacting_pair, 
                        m = t(selected_means[i,]), pval = t(selected_pvalues[i,]))
  temp_df <- dplyr::rename(temp_df, m = colnames(temp_df)[3])
  temp_df <- dplyr::rename(temp_df, pval = colnames(temp_df)[4])
  selected_df <- rbind(selected_df,temp_df)
}

#lr_interest <- unique(selected_df$gene_pair)
#setwd(wd)
#write.table(lr_interest, 'lr_interest.txt',sep = '\t', row.names = F, col.names = F)

setwd(wd)
tiff("GE.tiff", width=1100, height=1000, res=120) #change accordingly

f <- ggplot(selected_df, aes(x = cell_type, y = gene_pair)) +
  geom_point(aes(size = pval, colour = m)) + 
  theme(text = element_text(size=18), axis.text.x = element_text(angle = 20, hjust = 1)) +
  scale_color_gradient(low="blue", high="red") +
  xlab("Interacting cell types") +
  ylab("Receptor-ligand pairs") +
  theme(legend.position="top") +
  guides(size=guide_legend(title="P value (-log10)")) +
  guides(colour=guide_legend(title="Mean Expression (log2)"))
#  theme(panel.background = element_rect(fill = "white", colour = "white"), 
#        panel.grid.major = element_line(size = 0.5, linetype = 'solid',colour = "grey"))
f

dev.off()
```

```{r NicheNet setup}
#devtools::install_github("saeyslab/nichenetr")
library(nichenetr)
library(tidyverse)
library(ggplot2)
library(circlize)
library(dplyr)
library(Seurat)

# Load Seurat object and add meta.data for differential expression conditions
object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/GMHvsCTRL_mgendo/GMHvsCTRL_mgendo.rds")
object$celltype_copy <- object$celltype
object <- SetIdent(object, value = object[["celltype_copy"]]); levels(object)

## TEST ##
test = subset(object_og, idents = "Mitotic")
test <- SetIdent(test, value = test[["niche"]]); levels(test)
test_DEG = FindMarkers(test, ident.1 = "GeEndo", ident.2 = "CtxEndo", min.pct = 0.1)
sum(test_DEG$avg_log2FC > 0) # number of upreg genes in GE (ident.1)
sum(test_DEG$avg_log2FC < 0) # number of upreg genes in CTX (ident.2)

#subset Seurat obj accordingly
object <- subset(x = object_2023, idents = c("Mitotic", "1 Homeostatic MG 1","5 Vascul-assoc MG","6 Monocytes","7 HLA+ My")) #"Capillary" "Arterial"  "Venous"   "Tip"   "Mitotic"  
levels(x = object)

#colnames(x = object[[]])
object$celltype <- object$niche2 #need to specify celltype for downstream NicheNet analysis
object@meta.data$celltype %>% table() #should be categories in niche
object <- SetIdent(object, value = object[["niche2"]])
Idents(object = object) %>% table() #should be categories in niche

# Define groups (needed for downstream analysis)
new_label <- object@meta.data$niche2
new_label[new_label=="ENDO_ctrl"] <- "CTRL"
new_label[new_label=="MG_ctrl"] <- "CTRL"
new_label[new_label=="MG_GMH"] <- "GMH"
new_label[new_label=="ENDO_GMH"] <- "GMH"
object$groups <- new_label
object@meta.data$groups %>% table()
#DimPlot(object, reduction = "umap", group.by = "groups") 

# Load prior ligand-receptor model from NicheNet
# ref: https://zenodo.org/record/3260758 
ligand_target_matrix = readRDS("/Users/jiapei/Desktop/single_cell_analysis/NicheNet/ligand_target_matrix.rds")
# target genes in rows, ligands in columns
lr_network = readRDS("/Users/jiapei/Desktop/single_cell_analysis/NicheNet/lr_network.rds")
weighted_networks = readRDS("/Users/jiapei/Desktop/single_cell_analysis/NicheNet/weighted_networks.rds")
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
#head(weighted_networks$lr_sig) # interactions and their weights in the ligand-receptor + signaling network
#head(weighted_networks$gr) # interactions and their weights in the gene regulatory network
```
```{r NicheNet human genes to mouse genes}
# NicheNet genes are human, the following code is to convert to orthogonal mouse genes
lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()
ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]
weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```

```{r NicheNet define receivers, senders, DE comparison}
# Define sender/niche and receiver/target
levels(x = object)

## receiver
receiver = c("ENDO_ctrl","ENDO_GMH")
receiver_pct = 0.2
#expressed_genes_receiver = get_expressed_genes(receiver, object, pct = receiver_pct)
#length(expressed_genes_receiver) #genes expr by pct(%) cells
list_expressed_genes_receiver = receiver %>% unique() %>% lapply(get_expressed_genes, object, pct = receiver_pct) #, assay_oi = "SCT"
#set threshold at 0.8 to exclude any genes that might belong to sender cell type
expressed_genes_receiver = list_expressed_genes_receiver %>% unlist() %>% unique() #check size of vector! (not empty and ~1000)
background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]

## sender
sender_celltypes = c("MG_ctrl","MG_GMH")
sender_pct = 0.2
#expressed_genes_sender_check = get_expressed_genes(sender_celltypes, object, pct = sender_pct)
#expressed_genes_sender_check                                                               
list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, object, pct = sender_pct, assay_oi = "SCT")
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique() #check size of vector! (not empty and ~1000)
# remove ligands that are expressed on receiver population (for cell type specificity); comment out if all ligands get removed!
#nonsp = c(expressed_genes_receiver,expressed_genes_sender)
#nonsp = nonsp[duplicated(nonsp)]
#expressed_genes_sender = expressed_genes_sender[!expressed_genes_sender %in% nonsp]

# Define gene set of interest
seurat_obj_receiver= subset(object, idents = receiver)
seurat_obj_receiver = SetIdent(seurat_obj_receiver, value = seurat_obj_receiver[["groups"]])
condition_oi = "GMH"
condition_reference = "CTRL" # order doesn't matter, unless you specify only.pos = T
DE_table_receiver = FindMarkers(object = seurat_obj_receiver, ident.1 = condition_oi, ident.2 = condition_reference, 
                                min.pct = 0.1) %>% rownames_to_column("gene") # very few genes upreg in GE for endo (more downreg)
# only.pos = T <- enriched in condition_oi; but looks like ligand prediction ends up with the same set of genes regardless or pos or neg
# increased min.pct from 0.1 to 0.25 (for higher condition specificity)
setwd(wd)
write.table(DE_table_receiver, 'DE_table_receiver.txt',sep = '\t', row.names = F, col.names = T)

# no filter
geneset_oi = DE_table_receiver %>% pull(gene)
# or filter
geneset_oi = DE_table_receiver %>% filter(p_val_adj < 0.05 & avg_log2FC > 2) %>% pull(gene) # check DE_table_receiver to adjust avg_log2FC
# or define geneset_oi manually
geneset_oi <- c()

geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```

```{r NicheNet define and analyze ligands}
# Define potential ligands
ligands = lr_network %>% pull(from) %>% unique()
receptors = lr_network %>% pull(to) %>% unique()
expressed_ligands = intersect(ligands,expressed_genes_sender) #affected by prev code/settings
expressed_receptors = intersect(receptors,expressed_genes_receiver) #affected by prev code/settings
potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()

# Ligand activity analysis
# i.e. rank potential ligands based on the presence of their target genes in gene set of interest 
ligand_activities = predict_ligand_activities(geneset = geneset_oi, background_expressed_genes = background_expressed_genes, 
                                              ligand_target_matrix = ligand_target_matrix, 
                                              potential_ligands = potential_ligands)
view(ligand_activities)
#ligand_activities = ligand_activities %>% arrange(-pearson) %>% mutate(rank = base::rank(dplyr::desc(pearson)))
#ligand_activities = subset(ligand_activities, pearson > 0.01) 

# Pearson correlation ~ 0: for ligands expressed in sender cells, there is no enrichment of target genes in the geneset of interest compared to the background; i.e. no evidence that sender cell is regulating gene expression changes in receiver cell
# Pearson correlation < 0: there is enrichment of top predicted target genes in the background compared to the gene set of interest. When these values are negative, but very close to zero, there is nothing to worry about. But when these values are really negative, this can indicate that something went wrong in the analysis, probably in the definition of background and gene set of interest.
# If you get AUROC values < 0.5 and negative PCCs, this indicate that there are more targets of selected ligands in the background of expressed genes compared to DE genes between your conditions. This can be due to the possible false assumption that ligands from sender cells cause differences between conditions in receiver cell.

# only include ligands with positive pearson values to predict target genes
best_upstream_ligands = subset(ligand_activities, pearson > 0) %>% pull(test_ligand) %>% unique()
best_upstream_ligands <- best_upstream_ligands[!best_upstream_ligands %in% c("OSM")] # remove any unwanted ligands (confusing ones)
# top 20 ranked ligands that can be used to predict target genes
best_upstream_ligands = ligand_activities %>% top_n(20, pearson) %>% arrange(-pearson) %>% pull(test_ligand) %>% unique()
# or define it as your ligands of interest
best_upstream_ligands = ligand_activities$test_ligand #no filter
avg_expr = AverageExpression(object, features = best_upstream_ligands) 
avg_expr_df <- data.frame(unlist(avg_expr$RNA))
avg_expr_df <- avg_expr_df %>% mutate( diff = (CTX_MG + GE_MG)/2 / (CTX_ENDO + GE_ENDO)/2) # change accordingly
best_upstream_ligands <- rownames(subset(avg_expr_df, diff < 1/(2^50))) # pick genes that have higher expression in mg (change accordingly)
best_upstream_ligands
# or define it manually
best_upstream_ligands <- c("CXCL12","CX3CL1")

# Plot top ligands
setwd(wd)
tiff("Top_ligand.tiff", width=1800, height=600, res=120)
DotPlot(object, features = best_upstream_ligands %>% rev(), cols = "RdYlBu", assay = "RNA") + RotatedAxis()
dev.off()

# Plot top 20 ligands based on groups (e.g. region) and Violin plots of selected ligands 
#setwd(wd)
#tiff("MgEndo_select_ligand_mg_groups.tiff", width=1000, height=800, res=120)
#DotPlot(object, features = nichenet_output$top_ligands %>% rev(), split.by = "groups") + RotatedAxis()
#tiff("MgEndo_23gw_top20ligand_endo_genes.tiff", width=2000, height=400, res=120)
#VlnPlot(object = object, features = c("PECAM1", "ADAM17","ITGAM"), split.by = "groups", pt.size = 0, combine = T)
#dev.off()
```

```{r NicheNet infer TARGETS}
# Infer receptors
active_ligand_target_links_df = best_upstream_ligands %>% 
  lapply(get_weighted_ligand_target_links, geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>% 
  bind_rows() %>% drop_na() # n determines number of targets per ligand that are included in the downstream analysis
active_ligand_target_links = prepare_ligand_target_visualization(
  ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.1) #higher cutoff = more stringent

order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
order_targets = active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()

rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() 
colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() 

vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()

p_ligand_target_network = vis_ligand_target %>% 
  make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple", legend_position = "right", 
                      x_axis_position = "top", legend_title = "Regulatory potential")  + 
  theme(axis.text.x = element_text(face = "italic")) +
  theme(axis.text.x.top = element_text(vjust = 0.5)) + 
  theme(legend.direction="vertical") +
  scale_fill_gradient2(low = "whitesmoke", high = "purple")
# if ligands are left out, it's because they dont have target genes with high enough regulatory potential scores
# To include them, you can be less stringent in the cutoff
# To increase number of targets, increase n (see above)

# Plot predicted target genes of top ligands
setwd(wd)
tiff("Top_ligand_predictTarget.tiff", width=2000, height=400, res=120)
p_ligand_target_network
dev.off()
```

```{r NicheNet infer RECEPTORS}
# Find receptors of top-ranked ligands
lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()

lr_network_top_df_large = weighted_networks_lr %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors)

lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
lr_network_top_matrix = lr_network_top_df %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)

dist_receptors = dist(lr_network_top_matrix, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]
    
dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix))
order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix))

vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
rownames(vis_ligand_receptor_network) = order_receptors %>% make.names()
colnames(vis_ligand_receptor_network) = order_ligands_receptor %>% make.names()

p_ligand_receptor_network = vis_ligand_receptor_network %>% t() %>% 
  make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", 
                      x_axis_position = "top",legend_title = "Prior interaction potential") +
  theme(axis.text.x.top = element_text(vjust = 0.5))

# Plot predicted receptors of top 20 ligands
setwd(wd)
tiff("Top_ligand_predictReceptor.tiff", width=1500, height=1000, res=120)
p_ligand_receptor_network
dev.off()

# Further filtering potential ligand-receptor pairs with published literature and databses
lr_network_strict = lr_network %>% filter(database != "ppi_prediction_go" & database != "ppi_prediction")
ligands_bona_fide = lr_network_strict %>% pull(from) %>% unique()
receptors_bona_fide = lr_network_strict %>% pull(to) %>% unique()

lr_network_top_df_large_strict = lr_network_top_df_large %>% distinct(from,to) %>% inner_join(lr_network_strict, by = c("from","to")) %>% distinct(from,to)
lr_network_top_df_large_strict = lr_network_top_df_large_strict %>% inner_join(lr_network_top_df_large, by = c("from","to"))

lr_network_top_df_strict = lr_network_top_df_large_strict %>% spread("from","weight",fill = 0)
lr_network_top_matrix_strict = lr_network_top_df_strict %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df_strict$to)

dist_receptors = dist(lr_network_top_matrix_strict, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]

dist_ligands = dist(lr_network_top_matrix_strict %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix_strict))
order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix_strict))

vis_ligand_receptor_network_strict = lr_network_top_matrix_strict[order_receptors, order_ligands_receptor]
rownames(vis_ligand_receptor_network_strict) = order_receptors %>% make.names()
colnames(vis_ligand_receptor_network_strict) = order_ligands_receptor %>% make.names()

p_ligand_receptor_network_strict = vis_ligand_receptor_network_strict %>% t() %>% 
  make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", 
                      x_axis_position = "top",legend_title = "Prior interaction potential\n(bona fide)")

# Plot predicted receptors (published) of top 20 ligands
setwd(wd)
tiff("Top_ligand_verifiedReceptor.tiff", width=500, height=500, res=120)
p_ligand_receptor_network_strict
dev.off()
```

```{r NicheNet Circos plot - start with initialized ligand_target_matrix etc from prev NicheNet code}
# this code block allows more customization
#best_upstream_ligands <- best_upstream_ligands[!best_upstream_ligands %in% c("APP","APLN")] #exclude any ligands with no target genes
avg_expression_ligands = AverageExpression(object, features = best_upstream_ligands, group.by = "region", assays = "RNA")

# which cell subpop express top ligands? seems to include both receiver and sender cell subpops
sender_ligand_assignment = avg_expression_ligands$RNA %>% 
  apply(1, function(ligand_expression){ligand_expression > (ligand_expression %>% mean() + ligand_expression %>% sd())}) %>% t()
# remove + ligand_expression %>% sd() if sender_ligand_assigment is a matrix with all FALSE values
sender_ligand_assignment = sender_ligand_assignment %>% 
  apply(2, function(x){x[x == TRUE]}) %>% purrr::keep(function(x){length(x) > 0})

# which ligands are common across multiple cell types?
all_assigned_ligands = sender_ligand_assignment %>% lapply(function(x){names(x)}) %>% unlist()
unique_ligands = all_assigned_ligands %>% table() %>% .[. == 1] %>% names()
general_ligands = best_upstream_ligands %>% setdiff(unique_ligands)
sender_ligand_assignment

# change accordingly
#homeo_ligands = sender_ligand_assignment$`1 Homeostatic MG 1` %>% names() %>% setdiff(general_ligands)
#vam_ligands = sender_ligand_assignment$`5 Vascul-assoc MG` %>% names() %>% setdiff(general_ligands)
ctx_ligands = sender_ligand_assignment$`CTX` %>% names() %>% setdiff(general_ligands)
ge_ligands = sender_ligand_assignment$`GE` %>% names() %>% setdiff(general_ligands)
#remove_na = remove_na[!is.na(remove_na)] # remove any NA ligands

ligand_type_indication_df = tibble(
  ligand_type = c(rep("ctx", times = ctx_ligands %>% length()),
                  rep("ge", times = ge_ligands %>% length())), #rep("General", times = general_ligands %>% length())
  ligand = c(ctx_ligands, ge_ligands)) # change accordingly

# define ligand-targets of interest (n = 20: top 20 targets for each ligand)
#active_ligand_target_links_df = best_upstream_ligands %>% lapply(get_weighted_ligand_target_links, geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 2000) %>% bind_rows() %>% drop_na() #already ran in prev block
active_ligand_target_links_df = active_ligand_target_links_df %>% mutate(target_type = "up") %>% inner_join(ligand_type_indication_df)
active_ligand_target_links_df$ligand_type %>% table()

circos_links = active_ligand_target_links_df #check if empty, adjust prev parameters if so
```

```{r NicheNet Circos plot - ligands and TARGET genes}
# Circos visualization setup
grid_col_ligand =c("ctx" = "blue", "ge" = "red") # change accordingly # "General" = "beige"
grid_col_target =c("up" = "grey") # change accordingly (receiver color)

grid_col_tbl_ligand = tibble(ligand_type = grid_col_ligand %>% names(), color_ligand_type = grid_col_ligand)
grid_col_tbl_target = tibble(target_type = grid_col_target %>% names(), color_target_type = grid_col_target)

circos_links = circos_links %>% mutate(ligand = paste(ligand," ")) # extra space: differentiate btwn a gene as ligand and a gene as target!
circos_links = circos_links %>% inner_join(grid_col_tbl_ligand) %>% inner_join(grid_col_tbl_target)
links_circle = circos_links %>% select(ligand, target, weight)

ligand_color = circos_links %>% distinct(ligand,color_ligand_type)
grid_ligand_color = ligand_color$color_ligand_type %>% set_names(ligand_color$ligand)
target_color = circos_links %>% distinct(target,color_target_type)
grid_target_color = target_color$color_target_type %>% set_names(target_color$target)

grid_col =c(grid_ligand_color,grid_target_color)

# give the option that links in the circos plot will be transparant ~ ligand-target potential score
transparency = circos_links %>% mutate(weight =(weight-min(weight))/(max(weight)-min(weight))) %>% 
  mutate(transparency = 1-weight) %>% .$transparency 

# order ligands and targets
target_order = circos_links$target %>% unique()
# change accordingly!
set_ligand_order = c(ctx_ligands, ge_ligands) #general_ligands
ligand_order = set_ligand_order %>% c(paste(.," ")) %>% intersect(circos_links$ligand)
order = c(ligand_order,target_order)

# define gaps between different segments
width_same_cell_same_ligand_type = 0.5
width_different_cell = 0.5 #width_different_cell = 4
width_ligand_target = 8
width_same_cell_same_target_type = 0.5

# change accordingly
gaps_init = c(
  rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "ctx") %>% distinct(ligand) %>% nrow() -1)),
  width_different_cell,
  rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "ge") %>% distinct(ligand) %>% nrow() -1)),
  #width_different_cell,
  #rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "General") %>% distinct(ligand) %>% nrow() -1)),
  width_ligand_target)

gaps = c(gaps_init,
  rep(width_same_cell_same_target_type, times = (circos_links %>% filter(target_type == "up") %>% distinct(target) %>% nrow() -1)),
  width_ligand_target) # skipped activated and prolif mg because no ligand identified? 

library(ComplexHeatmap)
circos.par(gap.degree = gaps)
lgd_colors = c("blue","red") # to be reused in ligand-receptor circos plot
lgd_names = c("CTX CD45+ (ligands)", "GE CD45+ (ligands)") # to be reused in ligand-receptor circos plot # Shared (ligands)
lgd_lines = Legend(at = c(lgd_names,"Endothelial (targets)"), type = "points", 
  legend_gp = gpar(col = c(lgd_colors,"grey")), 
  title_position = "topleft", title = "Cell type")

setwd(wd)
tiff("TOPligand_target_circo.tiff", width=1200, height=1200, res=120)
# width proportional to ligand-target regulatory potential (based on prior knowledge)
chordDiagram(links_circle, directional = 1, order = order, link.sort = T, link.decreasing = F, 
             grid.col = grid_col, transparency = 0, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"), 
             link.arr.type = "big.arrow", annotationTrack = "grid", preAllocateTracks = list(track.height = 0.075))
# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
                                                                      facing = "clockwise", niceFacing = T, adj = c(0,0.55), cex =1)},
             bg.border = NA)
# add legend
draw(lgd_lines, x = unit(4, "mm"), y = unit(4, "mm"), just = c("left", "bottom"))
dev.off()

setwd(wd)
tiff("TOPligand_target_circo_transparent.tiff", width=1200, height=1200, res=120)
# transparency proportional to ligand-target regulatory potential (based on prior knowledge)
chordDiagram(links_circle, directional = 1, order = order,link.sort = T, link.decreasing = F, grid.col = grid_col,
             transparency = transparency, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"), link.arr.type = "big.arrow",
             annotationTrack = "grid", preAllocateTracks = list(track.height = 0.075))
# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,facing = "clockwise", niceFacing = TRUE, 
               adj = c(0, 0.55), cex = 1)}, bg.border = NA)
# add legend
draw(lgd_lines, x = unit(4, "mm"), y = unit(4, "mm"), just = c("left", "bottom"))
dev.off()
```

```{r NicheNet Circos plot - ligands and RECEPTORS}
lr_network_top_df = weighted_networks$lr_sig %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors) %>% 
  dplyr::rename(ligand = from, receptor = to)
lr_network_top_df = lr_network_top_df %>% mutate(receptor_type = "receptor") %>% inner_join(ligand_type_indication_df)

## uncomment following line if not using customized geneset (e.g. only.pos = T) ##
#lr_network_top_df = nichenet_output$ligand_receptor_df %>% mutate(receptor_type = "endo_receptor") %>% inner_join(ligand_type_indication_df)

#grid_col_ligand =c("General" = "lawngreen","ctrl_mg_specific" = "royalblue","GMH_mg_specific" = "red") #defined in prev block, uncomment if you didn't run the prev block
grid_col_receptor =c("receptor" = "grey") #"pink" "lightblue"

grid_col_tbl_ligand = tibble(ligand_type = grid_col_ligand %>% names(), color_ligand_type = grid_col_ligand)
grid_col_tbl_receptor = tibble(receptor_type = grid_col_receptor %>% names(), color_receptor_type = grid_col_receptor)

circos_links = lr_network_top_df %>% mutate(ligand = paste(ligand," ")) # extra space between a gene as ligand and a gene as receptor!
circos_links = circos_links %>% inner_join(grid_col_tbl_ligand) %>% inner_join(grid_col_tbl_receptor)
links_circle = circos_links %>% select(ligand, receptor, weight)

ligand_color = circos_links %>% distinct(ligand,color_ligand_type)
grid_ligand_color = ligand_color$color_ligand_type %>% set_names(ligand_color$ligand)
receptor_color = circos_links %>% distinct(receptor,color_receptor_type)
grid_receptor_color = receptor_color$color_receptor_type %>% set_names(receptor_color$receptor)

grid_col = c(grid_ligand_color,grid_receptor_color)

# give the option that links in the circos plot will be transparant ~ ligand-receptor potential score
transparency = circos_links %>% mutate(weight =(weight-min(weight))/(max(weight)-min(weight))) %>% mutate(transparency = 1-weight) %>% .$transparency 

receptor_order = circos_links$receptor %>% unique()
ligand_order = set_ligand_order %>% c(paste(.," ")) %>% intersect(circos_links$ligand)
order = c(ligand_order,receptor_order)

#width_same_cell_same_ligand_type = 0.5
#width_different_cell = 4
width_ligand_receptor = 8
width_same_cell_same_receptor_type = 0.5

# change accordingly
gaps_init = c(
  rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "ctx") %>% distinct(ligand) %>% nrow() -1)),
  width_different_cell,
  rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "ge") %>% distinct(ligand) %>% nrow() -1)),
  #width_different_cell,
  #rep(width_same_cell_same_ligand_type, times = (circos_links %>% filter(ligand_type == "General") %>% distinct(ligand) %>% nrow() -1)),
  width_ligand_target) #need to run gaps_init again since circos_links changed!

gaps = c(
  gaps_init,
  rep(width_same_cell_same_receptor_type, times = (circos_links %>% filter(receptor_type == "receptor") %>% distinct(receptor) %>% nrow() -1)),
  width_ligand_receptor)

circos.par(gap.degree = gaps)
lgd_lines = Legend(at = c(lgd_names,"Endothelial (receptors)"), type = "points", 
  legend_gp = gpar(col = c(lgd_colors,"grey")),
  title_position = "topleft", title = "Cell type")

setwd(wd)
tiff("TOPligand_circo_receptor.tiff", width=1200, height=1200, res=120)
chordDiagram(links_circle, directional = 1, order = order, link.sort = T, link.decreasing = F, 
             grid.col = grid_col,transparency = 0, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", annotationTrack = "grid", preAllocateTracks = list(track.height = 0.075)) 
            #link.visible = links_circle$weight >= cutoff_include_all_ligands
circos.track(track.index = 1, panel.fun = function(x, y) {circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
                                                                      facing = "clockwise", niceFacing = T, adj = c(0, 0.55), cex = 1)},
             bg.border = NA)
# add legend
draw(lgd_lines, x = unit(4, "mm"), y = unit(4, "mm"), just = c("left", "bottom"))
dev.off()

setwd(wd)
tiff("TOPligand_circo_receptor_transparent.tiff", width=1200, height=1200, res=120)
chordDiagram(links_circle, directional = 1, order = order, link.sort = T, link.decreasing = F, 
             grid.col = grid_col, transparency = transparency, diffHeight = 0.005, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow",annotationTrack = "grid", preAllocateTracks = list(track.height = 0.075))
            #link.visible = links_circle$weight >= cutoff_include_all_ligands
circos.track(track.index = 1, panel.fun = function(x, y) {circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
                                                                      facing = "clockwise", niceFacing = T, adj = c(0,0.55), cex = 1)},
             bg.border = NA) 
# add legend
draw(lgd_lines, x = unit(4, "mm"), y = unit(4, "mm"), just = c("left", "bottom"))
dev.off()
```

# Single cell RNA-seq analysis using Seurat
## Basic quality control and filtering
We start the analysis after two preliminary steps: 1) ambient RNA correction using `soupX`; 2) doublet detection using `scrublet`. (not used here)
If starting from typical `Cell Ranger` output, it's possible to choose if you want to use Ensemble ID or gene symbol for the count matrix. This is done using `gene.column` option; default is '2', which is gene symbol. 
Meta.data is the most important field for next steps. It can be accessed using both @ and [[]] operators. Right now it has 3 fields per celL: dataset ID, number of UMI reads detected per cell (nCount_RNA), and the number of expressed (detected) genes per same cell (nFeature_RNA). 

```{r cell cycle score}
library(Seurat)
library(ggplot2)
library(SingleR)
library(dplyr)
library(celldex)
library(RColorBrewer)
library(SingleCellExperiment)

srat <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/17_18_23_anchor_new/Mg_SCT.rds")
srat
#str(srat) #`str` commant allows us to see all fields of the class

#Let's calculate cell cycle scores (https://satijalab.org/seurat/archive/v3.1/cell_cycle_vignette.html). This has to be done after normalization and scaling. Seurat has a built-in list, cc.genes (older) and cc.genes.updated.2019 (newer), that defines genes involved in cell cycle. For CellRanger reference GRCh38 2.0.0 and above, use cc.genes.updated.2019 (three genes were renamed: MLF1IP, FAM64A and HN1 became CENPU, PICALM and JPT). For mouse cell cycle genes you can use the solution detailed [here](https://github.com/satijalab/seurat/issues/2493). 
cc.genes.updated.2019
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes

srat <- CellCycleScoring(srat, s.features = s.genes, g2m.features = g2m.genes)
table(srat[[]]$Phase)

VlnPlot(srat,features = c("S.Score","G2M.Score")) & theme(plot.title = element_text(size=10))
```

## Differential expression and marker selection
If some clusters lack any notable markers, adjust the clustering. It is [recommended](https://github.com/satijalab/seurat/discussions/4032) to do differential expression on the `RNA` assay, and not the `SCTransform`.

## Cell type annotation using SingleR
Given the markers that we've defined, we can mine the literature and identify each cell type. However, we can try automaic annotation with SingleR, which can be used with Seurat, SCE, etc. Detailed signleR manual can be found [here](http://bioconductor.org/books/release/SingleRBook/).

Let's get reference datasets from celldex package. Note that there are two cell type assignments, `label.main` and `label.fine`.

```{r cell type label 1, warning=F,message=F}
monaco.ref <- celldex::MonacoImmuneData()
hpca.ref <- celldex::HumanPrimaryCellAtlasData()
dice.ref <- celldex::DatabaseImmuneCellExpressionData()
```

Let's convert our Seurat object to single cell experiment (SCE) for convenience. After this, using `SingleR` becomes very easy: 

```{r cell type label 2}
sce <- as.SingleCellExperiment(DietSeurat(srat))
sce

monaco.main <- SingleR(test = sce,assay.type.test = 1,ref = monaco.ref,labels = monaco.ref$label.main)
monaco.fine <- SingleR(test = sce,assay.type.test = 1,ref = monaco.ref,labels = monaco.ref$label.fine)
hpca.main <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.main)
hpca.fine <- SingleR(test = sce,assay.type.test = 1,ref = hpca.ref,labels = hpca.ref$label.fine)
dice.main <- SingleR(test = sce,assay.type.test = 1,ref = dice.ref,labels = dice.ref$label.main)
dice.fine <- SingleR(test = sce,assay.type.test = 1,ref = dice.ref,labels = dice.ref$label.fine)
```

Let's see the summary of general cell type annotations. These match our expectations (and each other) reasonably well. 

```{r cell type label 3}
table(monaco.main$pruned.labels)
table(hpca.main$pruned.labels)
table(dice.main$pruned.labels)
#The finer cell types annotations are you after, the harder they are to get reliably. This is where comparing many databases, as well as using individual markers from literature, would all be very valuable.
table(monaco.fine$pruned.labels)
table(hpca.fine$pruned.labels)
table(dice.fine$pruned.labels)

#Let's add the annotations to the Seurat object metadata so we can use them: 
srat@meta.data$monaco.main <- monaco.main$pruned.labels
srat@meta.data$monaco.fine <- monaco.fine$pruned.labels
srat@meta.data$hpca.main   <- hpca.main$pruned.labels
srat@meta.data$dice.main   <- dice.main$pruned.labels
srat@meta.data$hpca.fine   <- hpca.fine$pruned.labels
srat@meta.data$dice.fine   <- dice.fine$pruned.labels

setwd(wd)
saveRDS(srat, file = "Mg_SCT.rds")
```

Finally, let's visualize the fine-grained annotations.

```{r cell type label 4}
srat <- SetIdent(srat, value = "dice.fine")
setwd(wd)
tiff("UMAP_dice_fine.tiff", width=1000, height=1000, res=120)
DimPlot(srat, label = T , repel = T, label.size = 3) + NoLegend()
dev.off()
```

A detailed book on how to do cell type assignment / label transfer with `singleR` [is available](https://bioconductor.org/books/3.12/SingleRBook/).

```{r ProjecTILs 1: set up reference dataset}
if (!requireNamespace("remotes")) install.packages("remotes")
library(remotes)

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if (!requireNamespace("Seurat", quietly = TRUE)) {
   install.packages("Seurat")
}

if (!requireNamespace("TILPRED", quietly = TRUE)) {
  install.packages(c("doParallel","doRNG"))
  BiocManager::install(c("AUCell","SingleCellExperiment"))
  remotes::install_github("carmonalab/TILPRED")
}

#remotes::install_github("carmonalab/ProjecTILs")
install.packages("/Users/jiapei/Desktop/single_cell_analysis/external_data/ProjecTILs", repos = NULL, type="source") #install from source
library(ProjecTILs)

# load reference object (onto which data will be projected)
#wd <- "/Users/jiapei/Desktop/single_cell_analysis/external_data"
#setwd(wd); ref <- load.reference.map() #sample data
ref_mg <- load.reference.map("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/mg_prcomp.rds")
#ref_mg <- load.reference.map("//Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/scvelo_mg_sample.rds")
DefaultAssay(ref_mg) <- "RNA"
ref_mg$functional.cluster <- ref_mg$celltype

# re-compute PCA embeddings outside of Seurat
set.seed(1234)
which.assay="integrated"
varfeat <- ref_mg@assays[[which.assay]]@var.features
refdata <- data.frame(t(ref_mg@assays[[which.assay]]@data[varfeat,]))
refdata <- refdata[, sort(colnames(refdata))]
ref.pca <- prcomp(refdata, rank. = 50, scale. = TRUE, center = TRUE, retx=TRUE)
ref.pca$rotation[1:5,1:5]
#ref.pca.seurat <- Loadings(object = ref_mg, reduction = "pca")

# re-compute UMAP outside of Seurat
library(umap)
seed=1234
n.neighbors=30
min.dist=0.2
metric="cosine"
ndim=15
umap.config <- umap.defaults
umap.config$n_neighbors = n.neighbors
umap.config$min_dist = min.dist
umap.config$metric = metric
umap.config$n_components = 2
umap.config$random_state = seed
umap.config$transform_state = seed
ref.umap <- umap(ref_mg@misc$pca_object$x[,1:ndim], config=umap.config)
colnames(ref.umap$layout) <- c("UMAP_1","UMAP_2")
ref.umap

# replace Seurat obj parameters (UMAP will be different...)
ref_mg@reductions$umap@cell.embeddings <- ref.umap$layout # don't change this? ###
ref.umap$layout <- ref_mg@reductions$umap@cell.embeddings # keep UMAP coordinates the same
ref_mg@reductions$pca@cell.embeddings <- ref.pca$x
ref_mg@reductions$pca@feature.loadings <- ref.pca$rotation
colnames(ref_mg@reductions$pca@cell.embeddings) <- gsub("PC(\\d+)", "PC_\\1", colnames(ref.pca$x), perl=TRUE)
colnames(ref_mg@reductions$pca@feature.loadings) <- gsub("PC(\\d+)", "PC_\\1", colnames(ref.pca$rotation), perl=TRUE)

#Store the complete PCA and UMAP object in @misc
ref_mg@misc$pca_object <- ref.pca
ref_mg@misc$umap_object <- ref.umap ###
ref_mg@misc$projecTILs="custom_atlas"

DimPlot(ref_mg, reduction = "umap", pt.size = 0.5, group.by = "celltype") # check new UMAP

wd <- "/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg" # save new object (prcomp and umap steps take long to run)
setwd(wd)
saveRDS(ref_mg, file = "mg_prcomp.rds")
```

```{r ProjecTILs 2: load query dataset and make projection}
# load query object
##Seurat object
querydata_mg <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/scvelo_mg_sample.rds")
#querydata_mg  <- subset(x = querydata_mg, subset = celltype == "8 Neuron-assoc MG")
DefaultAssay(querydata_mg) <- "RNA"

##10X format
fname2 <- "./sample_data"
querydata2 <- read.sc.query(fname2, type="10x")

##Raw count matrix from GEO
library(GEOquery)
geo_acc <- "GSE86028"
getGEOSuppFiles(geo_acc)
fname3 <- sprintf("%s/GSE86028_TILs_sc_wt_mtko.tpm.log2.txt.gz", geo_acc)
querydata3 <- read.sc.query(fname3, type = "raw.log2")

# convert mouse genes into human
musGenes <- rownames(querydata_mg)
# Basic function to convert mouse to human gene names
convertMouseGeneList <- function(x){
  require("biomaRt")
  human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
  mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
  genesV2 = getLDS(attributes = c("mgi_symbol"), filters = "mgi_symbol", values = x , mart = mouse, attributesL = c("hgnc_symbol"), 
                   martL = human, uniqueRows=T)
  #humanx <- unique(genesV2[, 2])
  return(genesV2)
}

MGI_HGNC <- convertMouseGeneList(musGenes)
#take out duplicated mouse and human genes
humanGenes_unique <- unique(MGI_HGNC$HGNC.symbol)
index_unique <- match(humanGenes_unique, MGI_HGNC$HGNC.symbol)
MGI_HGNC <- MGI_HGNC[index_unique,]

musGenes_unique <- unique(MGI_HGNC$MGI.symbol)
index_unique <- match(musGenes_unique, MGI_HGNC$MGI.symbol)
MGI_HGNC <- MGI_HGNC[index_unique,]

keep <- match(MGI_HGNC$MGI.symbol, musGenes)
#subset query data to keep features to mouse genes with equivalent human genes
counts <- GetAssayData(querydata_mg, assay = "RNA")
counts <- counts[keep,]
querydata_mg_human <- subset(querydata_mg, features = rownames(counts))
#modify subsetted query data's RNA assay with human gene symbols
RNA <- querydata_mg_human@assays$RNA
newname_index <- match(RNA@counts@Dimnames[[1]], MGI_HGNC$MGI.symbol)
newname <- MGI_HGNC[newname_index,]$HGNC.symbol
RNA@counts@Dimnames[[1]] <- newname
RNA@data@Dimnames[[1]] <- newname
querydata_mg_human@assays$RNA <- RNA
#RNA@counts@Dimnames[[1]] %>% head(3)
#n_occur <- data.frame(table(keep))
#n_occur<- n_occur[n_occur$Freq > 1,]

# make projection
#query.list <- SplitObject(querydata_mg, split.by = "orig.ident")
query.projected_mg <- make.projection(querydata_mg, ref = ref_mg, filter.cells = F, query.assay = "RNA", human.ortho = F) #don't filter for T cells; had to change raw code to make sure human.ortho = FALSE (otherwise "too many genes missing" error)

setwd(wd)
tiff("projection.tiff", width=1000, height=1000, res=120)
plot.projection(ref_mg, query.projected_mg, cols=colors)
dev.off()
```

```{r ProjecTILs 3: cell state prediction}
query.projected_mg <- cellstate.predict(ref=ref_mg, query=query.projected_mg)
table(query.projected_mg$functional.cluster)

setwd(wd)
tiff("cellstate_perc.tiff", width=800, height=800, res=120)
plot.statepred.composition(ref_mg, query.projected_mg, metric = "Percent") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
  #+ geom_bar(stat="identity", fill=colors)
dev.off()

genes4radar <- c("CX3CR1","CCL2","SPP1","GPNMB","MKI67","TOP2A","CLDN5","MFSD2A","LYZ","S100A9","STMN2","DCX")

setwd(wd)
tiff("cellstate_radar.tiff", width=1500, height=1500, res=120)
plot.states.radar(ref_mg, query=query.projected_mg, min.cells = 20, genes4radar = genes4radar)
dev.off()
```

## Weighted gene coexpression network analysis (WGCNA)
https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/  
```{r WGCNA 1: load, clean datasets}
#BiocManager::install("WGCNA")
library(WGCNA)
options(stringsAsFactors = FALSE)

wd <- "/Users/jiapei/Desktop/single_cell_analysis/WGCNA"
#example data set
setwd(wd)
femData <- read.csv("FemaleLiver_Data/LiverFemale3600.csv")
dim(femData)
names(femData)

# Remove excess data
datExpr0 = as.data.frame(t(femData[, -c(1:8)]))
names(datExpr0) = femData$substanceBXH
rownames(datExpr0) = names(femData)[-c(1:8)]
# remove genes and samples with too many missing values
gsg = goodSamplesGenes(datExpr0, verbose = 3)
gsg$allOK
# if prev statement did not return TRUE:
if (!gsg$allOK)
{# print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0)
  printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
  printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}

# cluster samples and find outlier
sampleTree = hclust(dist(datExpr0), method = "average");
sizeGrWindow(12,9)
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
abline(h = 15, col = "red")# Plot a line to show the cut (i.e. threshold for outlier cutoff, user can arbitrarily define)
# Determine the outlier sample
clust = cutreeStatic(sampleTree, cutHeight = 15, minSize = 10)
table(clust)
# refine dataset to exclude the outlier
keepSamples = (clust==1)
datExpr = datExpr0[keepSamples, ]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)

# input clinical data on each sample (similar to sample sheet in bulk RNAseq analyses)
setwd(wd)
traitData = read.csv("FemaleLiver_Data/ClinicalTraits.csv")
dim(traitData)
names(traitData)
# remove columns that hold information we do not need.
allTraits = traitData[, -c(31, 16)]
allTraits = allTraits[, c(2, 11:36) ]
dim(allTraits)
names(allTraits)
# Form a data frame analogous to expression data that will hold the clinical traits.
femaleSamples = rownames(datExpr)
traitRows = match(femaleSamples, allTraits$Mice)
datTraits = allTraits[traitRows, -1]
rownames(datTraits) = allTraits[traitRows, 1]
collectGarbage()

# expression data is now in datExpr, and the corresponding clinical traits in datTraits
# Re-cluster samples
sampleTree2 = hclust(dist(datExpr), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(datTraits, signed = FALSE);
# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors, groupLabels = names(datTraits), main = "Sample dendrogram and trait heatmap")

# save 
setwd(wd)
save(datExpr, datTraits, file = "FemaleLiver-01-dataInput.RData")
```

```{r WGCNA 2: network construction and module detection}
options(stringsAsFactors = FALSE)

## Method 1: 1-step

# Load the data saved in the first part
setwd(wd)
lnames = load(file = "FemaleLiver-01-dataInput.RData")
#The variable lnames contains the names of loaded variables
lnames

# Define a range of values for soft-thresholding power (important parameter)
powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2))
cex1 = 0.9
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], 
     xlab="Soft Threshold (power)", ylab="Scale Free Topology Model Fit,signed R^2", type="n", main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers, cex=cex1, col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90, col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab="Soft Threshold (power)", ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
# pick soft-threshold power of 6 (scale free topology model result flattens out after 6, see first graph)

# one-step function
setwd(wd)
net = blockwiseModules(datExpr, power = 6, TOMType = "unsigned", minModuleSize = 30, reassignThreshold = 0, mergeCutHeight = 0.25, 
                       numericLabels = T, pamRespectsDendro = F, saveTOMs = T, saveTOMFileBase = "femaleMouseTOM", verbose = 3)
# paramter settings:
  #relatively large minimum module size of 30; mergeCutHeight is threshold for module merging; function returns numeric, rather than color, labels for modules; save Topological Overlap Matrix (TOM) = T
# function output:
  # net$colors contains module assignment; net$MEs contains module eigengenes
# We encourage the user to read the help file and play with parameter values. 
# maxBlockSize: size of largest block that the computer can handle. Default value: 5000. If dataset > 5000 probes, blockwiseModules will split dataset into several blocks. This will break some of the plotting code below. 
# For larger datasets: increase maxBlockSize
  #16GB workstation can handle 20000 probes; 32GB workstation can handle 30000. 4GB can handle 8000-10000

#To see how many modules were identified and what the module sizes are:
table(net$colors) #in order of descending size; label 0 is reserved for genes outside of all modules

# open a graphics window
sizeGrWindow(12, 9)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]], "Module colors", dendroLabels = F, hang = 0.03,
                    addGuide = T, guideHang = 0.05)
# function recutBlockwiseTrees can recut the dendrogram tree without recomputing the network and clustering

# save the module assignment and module eigengene information
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs
geneTree = net$dendrograms[[1]]
setwd(wd)
save(MEs, moduleLabels, moduleColors, geneTree, file = "FemaleLiver-02-networkConstruction-auto.RData")

## Method 2: step-by-step

# first run the soft-thresholding power block from method 1 (see above) -> select appropriate soft-threshold value
softPower = 6
adjacency = adjacency(datExpr, power = softPower)
# Turn adjacency into topological overlap
TOM = TOMsimilarity(adjacency)
dissTOM = 1-TOM
# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average")
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", labels = FALSE, hang = 0.04)

# There are many methods for branch cutting; our standard method is the Dynamic Tree Cut (dynamicTreeCut)
# We like large modules, so we set minimum module size relatively high:
minModuleSize = 30
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, deepSplit = 2, pamRespectsDendro = F, minClusterSize = minModuleSize)
table(dynamicMods) # list modules from largest to smallest size; module 0 are genes outside of all modules
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)

# Plot the dendrogram and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut", dendroLabels = F, hang = 0.03, addGuide = T, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")

# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs)
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average")
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes", xlab = "", sub = "")
# choose height cut of 0.25 (= correlation of 0.75)
MEDissThres = 0.25
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")

# Call an automatic merging function
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs
# plot dendrogram with original and merged modules underneath
sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors), c("Dynamic Tree Cut", "Merged dynamic"), dendroLabels = F, 
                    hang = 0.03, addGuide = T, guideHang = 0.05)
#dev.off()

# save data (with merged modules)
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50))
moduleLabels = match(moduleColors, colorOrder)-1
MEs = mergedMEs
# Save module colors and labels for use in subsequent parts
setwd(wd)
save(MEs, moduleLabels, moduleColors, geneTree, file = "FemaleLiver-02-networkConstruction-stepByStep.RData")

## Method 3: block-wise, for large datasets (although preferable: analyze dataset in 1 block)

# first run the soft-thresholding power block from method 1 (see above) -> select appropriate soft-threshold value
# Let's assume that hardware restricts us to analyze 2000 genes at one time. The basic idea is to use a two-level clustering. 
  # First, we use a fast and crude clustering method to pre-cluster genes into size ~ 2000 genes. 
  # We then perform a full network analysis in each block separately. 
  # At the end, modules whose eigengenes are highly correlated are merged. 
# Advantages: smaller memory req and faster calculations
# Trade-off: modules may not be optimal
# blockwiseModules can split datasets into blocks automatically; just need to specify largest number of genes to fit in a block:
setwd(wd)
bwnet = blockwiseModules(datExpr, maxBlockSize = 2000, power = 6, TOMType = "unsigned", minModuleSize = 30, reassignThreshold = 0,
                         mergeCutHeight = 0.25, numericLabels = T, saveTOMs = T, saveTOMFileBase = "femaleMouseTOM-blockwise",
                         verbose = 3)
# Relabel block-wise module labels so that modules with significant overlap with single-block modules have the same label:
# Load the results of single-block analysis
setwd(wd)
load(file = "FemaleLiver-02-networkConstruction-auto.RData")
# Relabel blockwise modules
bwLabels = matchLabels(bwnet$colors, moduleLabels)
# Convert labels to colors for plotting
bwModuleColors = labels2colors(bwLabels)
table(bwLabels)

# Graph dendrogram
sizeGrWindow(6,6)
# Plot the dendrogram and the module colors underneath for block 1
plotDendroAndColors(bwnet$dendrograms[[1]], bwModuleColors[bwnet$blockGenes[[1]]], "Module colors", 
                    main = "Gene dendrogram and module colors in block 1", dendroLabels = F, hang = 0.03, addGuide = T, guideHang = 0.05)
# Plot the dendrogram and the module colors underneath for block 2
plotDendroAndColors(bwnet$dendrograms[[2]], bwModuleColors[bwnet$blockGenes[[2]]], "Module colors", 
                    main = "Gene dendrogram and module colors in block 2", dendroLabels = F, hang = 0.03, addGuide = T, guideHang = 0.05)

# Compare results of block-wise analysis to the results from method 1 (all data analyzed in a single block(. A simple visual check:
sizeGrWindow(12,9)
plotDendroAndColors(geneTree, cbind(moduleColors, bwModuleColors), c("Single block", "2 blocks"),
                    main = "Single block gene dendrogram and module colors", dendroLabels = F, hang = 0.03, 
                    addGuide = T, guideHang = 0.05)

# Verify that module eigengenes in the single-block and block-wise approaches are similar
# We first calculate module eigengenes based on the single block and block-wise module colors:
singleBlockMEs = moduleEigengenes(datExpr, moduleColors)$eigengenes
blockwiseMEs = moduleEigengenes(datExpr, bwModuleColors)$eigengenes
# Next we match the single-block and block-wise eigengenes by name and calculate the correlations of the corresponding eigengenes:
single2blockwise = match(names(singleBlockMEs), names(blockwiseMEs))
signif(diag(cor(blockwiseMEs[, single2blockwise], singleBlockMEs)), 3)
```

```{r WGCNA 3: add external info and ID genes}
# The following setting is important, do not omit
options(stringsAsFactors = FALSE)
# Load the expression and trait data saved in the first part
setwd(wd)
lnames = load(file = "FemaleLiver-01-dataInput.RData")
#The variable lnames contains the names of loaded variables.
lnames
# Load network data saved in the second part
setwd(wd)
lnames = load(file = "FemaleLiver-02-networkConstruction-auto.RData");
lnames

#Identify modules sig associated with traits: correlate eigengenes with external traits and look for the most significant ones:
# Define numbers of genes and samples
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
# Color code each association by correlation value:
sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(", signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3))
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor, xLabels = names(datTraits), yLabels = names(MEs), ySymbols = names(MEs), colorLabels = F,
               colors = greenWhiteRed(50), textMatrix = textMatrix, setStdMargins = F, cex.text = 0.5, zlim = c(-1,1),
               main = paste("Module-trait relationships"))

# Quantify associations of individual genes with our trait of interest (weight) by defining Gene Significance (GS) as absolute value of correlation between gene and trait
# For each module, we also define module membership (MM) as correlation of module eigengene and gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.
# Define variable weight containing the weight column of datTrait
weight = as.data.frame(datTraits$weight_g)
names(weight) = "weight"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"))
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))
names(geneModuleMembership) = paste("MM", modNames, sep="")
names(MMPvalue) = paste("p.MM", modNames, sep="")
geneTraitSignificance = as.data.frame(cor(datExpr, weight, use = "p"))
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))
names(geneTraitSignificance) = paste("GS.", names(weight), sep="")
names(GSPvalue) = paste("p.GS.", names(weight), sep="")

# Using GS and MM, we can identify genes that have high significance for weight as well as high MM in interesting modules. As an example, we look at the brown module that has the highest association with weight. We plot a scatterplot of GS vs. MM in the brown module:
module = "brown"
column = match(module, modNames)
moduleGenes = moduleColors==module
sizeGrWindow(7, 7)
par(mfrow = c(1,1))
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]), abs(geneTraitSignificance[moduleGenes, 1]), 
                   xlab = paste("Module Membership in", module, "module"), ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"), cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, 
                   col = module)
# GS and MM are highly correlated: genes highly associated with a trait = most important elements of modules associated with that trait

# Merge statistical information with gene annotation and write out a file that summarizes the most important results and can be inspected in Excel. Our expression data are only annotated by probe ID names:
names(datExpr) # return all probe IDs included in the analysis
names(datExpr)[moduleColors=="brown"] #return probe IDs belonging to the brown module. 
# Load probe annotation file provided by the manufacturer of expression arrays to connect probe IDs to gene names and Entrez codes
setwd("/Users/jiapei/Desktop/single_cell_analysis/WGCNA/FemaleLiver_Data")
annot = read.csv(file = "GeneAnnotation.csv")
dim(annot)
names(annot)
probes = names(datExpr)
probes2annot = match(probes, annot$substanceBXH)
# The following is the number or probes without annotation:
sum(is.na(probes2annot)) # Should return 0

# Create data frame with: probe ID, gene symbol, Entrez code, module color, GS for weight, MM and p-values 
# Modules will be ordered by their significance for weight, with the most significant ones to the left
geneInfo0 = data.frame(substanceBXH = probes, geneSymbol = annot$gene_symbol[probes2annot],
                       LocusLinkID = annot$LocusLinkID[probes2annot], moduleColor = moduleColors, geneTraitSignificance, GSPvalue)
modOrder = order(-abs(cor(MEs, weight, use = "p")))
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership))
{
  oldNames = names(geneInfo0)
  geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]], MMPvalue[, modOrder[mod]])
  names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""), paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
# Order the genes in geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.weight))
geneInfo = geneInfo0[geneOrder, ]

# save
setwd(wd)
write.csv(geneInfo, file = "geneInfo.csv")
# inspect it directly within R using the command fix(geneInfo).
```

```{r WGCNA 4: GO}
# The following setting is important, do not omit
options(stringsAsFactors = FALSE)
# Load the expression and trait data saved in the first part
lnames = load(file = "FemaleLiver-01-dataInput.RData")
#The variable lnames contains the names of loaded variables.
lnames
# Load network data saved in the second part
lnames = load(file = "FemaleLiver-02-networkConstruction-auto.RData");
lnames

# Export gene identifiers for GO and functional enrichment analyses, e.g. we output Entrez codes from brown module into a file:
# Read in probe annotation
setwd("/Users/jiapei/Desktop/single_cell_analysis/WGCNA/FemaleLiver_Data")
annot = read.csv(file = "GeneAnnotation.csv")
# Match probes in the data set to the probe IDs in the annotation file
probes = names(datExpr)
probes2annot = match(probes, annot$substanceBXH)
# Get the corresponding Locuis Link IDs (Entrez codes)
allLLIDs = annot$LocusLinkID[probes2annot]
# Choose interesting modules
intModules = c("brown", "red", "salmon")
for (module in intModules)
{
  # Select module probes
  modGenes = (moduleColors==module)
  # Get their entrez ID codes
  modLLIDs = allLLIDs[modGenes]
  # Write them into a file
  fileName = paste("LocusLinkIDs-", module, ".txt", sep="")
  setwd(wd)
  write.table(as.data.frame(modLLIDs), file = fileName, row.names = F, col.names = F)
}
# As background in the enrichment analysis, we will use all probes in the analysis
fileName = paste("LocusLinkIDs-all.txt", sep="")
setwd(wd)
write.table(as.data.frame(allLLIDs), file = fileName, row.names = F, col.names = F)

# WGCNA can perform GO enrichment analysis
# Calling the GO enrichment analysis function GOenrichmentAnalysis is very simple. The function takes a vector of module labels, and the Entrez (a.k.a. Locus Link) codes for the genes whose labels are given.
if (!requireNamespace("BiocManager", quietly = T))
    install.packages("BiocManager")
BiocManager::install("GO.db")
if (!requireNamespace("BiocManager", quietly = T))
    install.packages("BiocManager")
BiocManager::install("AnnotationDBI")
if (!requireNamespace("BiocManager", quietly = T))
    install.packages("BiocManager")
BiocManager::install("org.Hs.eg.db")
library("GO.db")
library("AnnotationDBI")
library("org.Hs.eg.db")

GOenr = GOenrichmentAnalysis(moduleColors, allLLIDs, organism = "mouse", nBestP = 10)
tab = GOenr$bestPTerms[[4]]$enrichment # enrichment table containing the 10 best terms for each module present in moduleColors. Names of the columns within the table can be accessed by
names(tab)
```

```{r slingshot}
#if (!requireNamespace("BiocManager", quietly=TRUE))
#    install.packages("BiocManager")
#BiocManager::install("kstreet13/slingshot")
#BiocManager::install("SingleCellExperiment")

library(slingshot)
library(SingleCellExperiment)

object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/scvelo_mg_sample.rds")
count_raw_2 <- as.matrix(GetAssayData(object, slot = "counts"))
#head(colnames(count_raw_2),3)
sce <- SingleCellExperiment(assays = List(counts = count_raw_2)) #count_raw imported from Seurat object of choice
rm(object) 
rm(count_raw) #save space

# gene filtering
geneFilter <- apply(assays(sce)$counts,1,function(x){
    sum(x >= 3) >= 10
}) # filtering to include only genes with > 3 counts in more than 10 cells
sce <- sce[geneFilter, ]

# normalization (does NOT include batch correction, is basically full quantile normalization to ensure same distribution of expression values for each cell)
FQnorm <- function(counts){
    rk <- apply(counts,2,rank,ties.method='min')
    counts.sort <- apply(counts,2,sort)
    refdist <- apply(counts.sort,1,median)
    norm <- apply(rk,2,function(r){ refdist[r] })
    rownames(norm) <- rownames(counts)
    return(norm)
}
assays(sce)$norm <- FQnorm(assays(sce)$counts)

# save SCE object
setwd(wd)
saveRDS(sce, file = "slingshot_sce.rds")

# dimension reduction (PCA)
#sce <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_Mg/by_sample/slingshot/slingshot_sce.rds")
pca <- prcomp(t(log1p(assays(sce)$norm)), scale. = F)
rd1 <- pca$x[,1:2]
plot(rd1, col = rgb(0,0,0,.5), pch=16, asp = 1)

# dimension reduction (UMAP)
library(uwot)
```

```{r cellchat}
# Source: https://htmlpreview.github.io/?https://github.com/sqjin/CellChat/blob/master/tutorial/CellChat-vignette.html 
## Install and read packages ##
#devtools::install_github("sqjin/CellChat")
library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)

## Input Data ##
# Gene expression of cells: genes should be in rows with rownames and cells in columns with colnames. Normalized data (e.g., library-size normalization and then log-transformed with a pseudocount of 1) is required as input for CellChat analysis. If user provides count data, we provide a normalizeData function to account for library size and then do log-transformed.
# User assigned cell labels: a data frame (rows = cells) consisting of cell labels, which will be used for defining cell groups.
object <- readRDS("/Users/jiapei/Desktop/single_cell_analysis/ctrl_MgEndo/MgEndo.rds")
data.input <- GetAssayData(object, slot = "data", assay = "SCT") # normalized data matrix
Idents(object) <- "cell_type"
meta <- data.frame(labels = Idents(object), row.names = names(Idents(object))) # manually create a dataframe with cell labels
unique(meta$labels) # check cell labels

## Create CellChat object ##
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")
cellchat
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group

## Set ligand-receptor database ##
# CellChatDB: manually curated database of literature-supported ligand-receptor interactions in human and mouse. CellChatDB in mouse contains 2,021 interactions: 60% secrete autocrine/paracrine, 21% ECM-receptor and 19% cell-cell contact CellChatDB in human contains 1,939 interactions: 62% paracrine/autocrine, 22% ECM-receptor and 17% cell-cell contact.
# Users can update CellChatDB by adding their own curated ligand-receptor pairs. Please check our tutorial on how to do it.
showDatabaseCategory(CellChatDB.human) # CellChatDB.mouse 
# Show the structure of the database
dplyr::glimpse(CellChatDB.human$interaction)
#cellchat@DB <- subsetDB(CellChatDB.human, search = "Secreted Signaling") # use a subset of CellChatDB for cell-cell communication analysis
cellchat@DB <- CellChatDB.human # set the used database in the object

## Preprocessing expression data ##
# We find over-expressed ligands or receptors in one cell group and then find over-expressed ligand-receptor interactions if either ligand or receptor is over-expressed.
# We also provide a function to project gene expression data onto protein-protein interaction (PPI) network. Specifically, a diffusion process is used to smooth genes expression values based on their neighbors defined in a high-confidence experimentally validated protein-protein network. This function is useful when analyzing single-cell data with shallow sequencing depth because the projection reduces the dropout effects of signaling genes, in particular for possible zero expression of subunits of ligands/receptors. USERS can skip this step and set raw.use = TRUE in the function computeCommunProb().
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multiprocess", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
 
# project gene expression data onto PPI (Optional: when running it, USER should set `raw.use = FALSE` in the function `computeCommunProb()` in order to use the projected data)
cellchat <- projectData(cellchat, PPI.human)

## Compute communication probability and infer communication network ##
# CellChat infers biologically significant cell-cell communication by assigning each interaction with a probability value and peforming a permutation test. CellChat models the probability of cell-cell communication by integrating gene expression with spatial locations as well as prior known knowledge of the interactions between signaling ligands, receptors and their cofactors using the law of mass action.

# In computeCommunProb, users can use different methods to calculate average gene expression. trimean default: 25% truncated mean (average gene expression is 0 if <25% cells in a group express this gene). To use 10% truncated mean: type = "truncatedMean" and trim = 0.1. 
# The function computeAveExpr can check average expression of genes of interest, e.g, computeAveExpr(cellchat, features = c("CXCL12","CXCR4"), type =  "truncatedMean", trim = 0.1).
# If well-known signaling pathways are not predicted, USER can lower trim value.

# To quickly examine the inference results, USER can set nboot = 20. i.e. pvalue < 0.05 means none of the permutation results are larger than the observed communication probability.

? computeCommunProb
cellchat <- computeCommunProb(cellchat, type = "truncatedMean", trim = 0.1, raw.use = T)
# options(future.globals.maxSize = 1000 * 1024^2) # error debugging; see: https://satijalab.org/seurat/archive/v3.0/future_vignette.html
# raw.use: use raw data ('object@data.signaling') or projected data ('object@data.project'). Set raw.use = F to use projected data when analyzing single-cell data with shallow sequencing depth because projected data reduce dropout effects of signaling genes.

## Filter out interactions in few cells ##
cellchat <- filterCommunication(cellchat, min.cells = 10)

## Extract inferred cellular communication network as a data frame ##
df.net <- subsetCommunication(cellchat, sources.use = c("6 HLA+ My"), targets.use = c("Mitotic","Venous","Arterial","Capillary","Tip"),
                              slot.name = "netP") # slot.name = "net" (LR pairs), signaling = c("TGFb")

## Infer cell-cell communication at signaling pathway level ##
# Inferred communication network of ligand-receptor pairs and signaling pathways is stored in the slot net and netP, respectively.
cellchat <- computeCommunProbPathway(cellchat)

## Calculate aggregated cell-cell communication network ##
# by counting number of links or summarizing communication probability. 
cellchat <- aggregateNet(cellchat) #sources.use = c("6 HLA+ My"), targets.use = c("Mitotic")

## Visualize cell-cell communication ##
# Visualize: number of interactions or total interaction strength (weights) between any two cell groups using circle plot.
setwd(wd)
tiff("CellChat_NumInteraction.tiff", width=1200, height=1200, res=120)
netVisual_circle(cellchat@net$count, vertex.weight = rowSums(cellchat@net$count), weight.scale = T, label.edge= F, 
                 title.name = "Number of interactions")
tiff("CellChat_Weight.tiff", width=1200, height=1200, res=120)
netVisual_circle(cellchat@net$weight, vertex.weight = rowSums(cellchat@net$weight), weight.scale = T, label.edge= F, 
                 title.name = "Interaction weights/strength")
dev.off()

# Examine signaling sent from each cell group. Here we control edge.weight.max so that we can compare edge weights between different networks.
mat <- cellchat@net$weight
setwd(wd)
tiff("CellChat_celltype.tiff", width=2000, height=2000, res=120)
par(mfrow = c(4,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
dev.off()

# Chord diagram: CellChat provides two functions netVisual_chord_cell and netVisual_chord_gene for visualizing cell-cell communication with different purposes and different levels. netVisual_chord_cell is used for visualizing the cell-cell communication between different cell groups (where each sector in the chord diagram is a cell group), and netVisual_chord_gene is used for visualizing the cell-cell communication mediated by mutiple ligand-receptors or signaling pathways (where each sector in the chord diagram is a ligand, receptor or signaling pathway.)

# Explnations of edge color/weight, node color/size/shape: In all visualization plots, edge colors are consistent with the sources as sender, and edge weights are proportional to the interaction strength. Thicker edge line indicates a stronger signal. In the Hierarchy plot and Circle plot, circle sizes are proportional to the number of cells in each cell group. In the hierarchy plot, solid and open circles represent source and target, respectively. In the Chord diagram, the inner thinner bar colors represent the targets that receive signal from the corresponding outer bar. The inner bar size is proportional to the signal strength received by the targets. Such inner bar is helpful for interpreting the complex chord diagram. Note that there exist some inner bars without any chord for some cell groups, please just igore it because this is an issue that has not been addressed by circlize package.

# Visualize network of signaling pathways using netVisual_aggregate, and visualize networks of LR pairs using netVisual_individual.
cellchat@netP$pathways # All signaling pathways showing significant communications

pathways.show <- c("CXCL") 
# Hierarchy plot
netVisual_aggregate(cellchat, signaling = c("IGF"), vertex.receiver = seq(1,5))
# vertex.receiver: numeric vector giving the index of cell groups as targets in the left part of plot. This plot has two components: the left portion shows autocrine and paracrine signaling to cell groups of interest (i.e. vertex.receiver), and the right portion shows autocrine and paracrine signaling to the remaining cell groups. <- seems to just reorder cell groups, otherwise the same as circle plot (see below)

# Circle Plot
setwd(wd)
tiff("CellChat_IL6.tiff", width=800, height=800, res=120)
netVisual_aggregate(cellchat, signaling = c("IL6"), layout = "circle")
dev.off()

# Circle Plot for significant LR pair
pairLR <- extractEnrichedLR(cellchat, signaling = c("CXCL"), geneLR.return = F) # extract all significant LR pairs and related genes
LR.show <- pairLR[8,]; LR.show # show one ligand-receptor pair
netVisual_individual(cellchat, signaling = c("CXCL"), pairLR.use = LR.show, layout = "circle") # layout = "chord"

# Chord diagram
setwd(wd)
tiff("CellChat_IGF_chord.tiff", width=1000, height=1000, res=120)
netVisual_aggregate(cellchat, signaling = c("IGF"), layout = "chord")
dev.off()

# Grouped Chord diagram
# netVisual_chord_cell: can adjust different parameters in the circlize package. For example, we can group cell clusters into different groups
levels(cellchat@idents)
group.cellType <- c(rep("ENDO", 5), rep("CD45", 11)) # grouping cell clusters into endo or CD45+ cells
names(group.cellType) <- levels(cellchat@idents)

setwd(wd)
tiff("CellChat_IGF_chord_celltype.tiff", width=1000, height=1000, res=120)
netVisual_chord_cell(cellchat, signaling = c("IGF"), group = group.cellType, title.name = paste0(pathways.show, " signaling network"))
dev.off()

# LR pair Chord diagram
levels(cellchat@idents)
setwd(wd)
tiff("CellChat_chord_IGF_VEGF.tiff", width=1500, height=1500, res=120)
netVisual_chord_gene(cellchat, sources.use = c(13,15,16), targets.use = c(1:5), signaling = c("IGF","VEGF"), 
                     lab.cex = 1, legend.pos.y = 30)
dev.off()

# Pathway Chord diagram
# show all the significant signaling pathways
setwd(wd)
tiff("CellChat_chord_pathway.tiff", width=1500, height=1500, res=120)
netVisual_chord_gene(cellchat, sources.use = c(13,15,16), targets.use = c(1:5), slot.name = "netP", legend.pos.x = 10)
dev.off()

# Heatmap
setwd(wd)
tiff("CellChat_VEGF_heatmap.tiff", width=1000, height=500, res=120)
netVisual_heatmap(cellchat, signaling = c("VEGF"), color.heatmap = "Reds")
dev.off()

# Contribution of LR pair to the overall signaling pathway
netAnalysis_contribution(cellchat, signaling = c("CXCL"))

# Violin plots
setwd(wd)
tiff("CellChat_VlnPlot_VEGF.tiff", width=1000, height=800, res=120)
plotGeneExpression(cellchat, signaling = "VEGF") # enriched.only = FALSE to show all genes
dev.off()

# Save ALL plots
# Access all the signaling pathways showing significant communications
pathways.show.all <- cellchat@netP$pathways
for (i in 1:length(pathways.show.all)) {
  # Visualize communication network associated with both signaling pathway and individual L-R pairs
  netVisual(cellchat, signaling = pathways.show.all[i], layout = "circle")
  # Compute and visualize the contribution of each ligand-receptor pair to the overall signaling pathway
  gg <- netAnalysis_contribution(cellchat, signaling = pathways.show.all[i])
  ggsave(filename=paste0(pathways.show.all[i], "_L-R_contribution.pdf"), plot=gg, width = 10, height = 10, units = 'in', dpi = 300)
}

# Visualize cell-cell communication by multiple LR or signaling pathways
# Bubble plot
# show all significant LR pairs from some cell groups ('sources.use') to other cell groups ('targets.use')
setwd(wd)
tiff("CellChat_bubble_IGF_VEGF.tiff", width=1000, height=500, res=120)
netVisual_bubble(cellchat, sources.use = c(1:5), targets.use = c(6:16), signaling = c("IGF","VEGF"), remove.isolate = FALSE) # similar to cellphoneDB heatmap
dev.off()

## Systems analysis of cell-cell communication network ##
# It can determine signaling sources, targets, mediators and influencers within a network using centrality measures from network analysis
# It can predict incoming and outgoing signals for specific cell types as well as coordinated responses among different cell types by leveraging pattern recognition approaches.
# It can group signaling pathways by similarity measures and performing manifold learning from functional and topological perspectives.
# It can delineate conserved and context-specific signaling pathways by joint manifold learning of multiple networks.

# Compute the network centrality scores
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # 'netP': network of signaling pathways
# Visualize centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(cellchat, signaling = c("IGF"), width = 8, height = 2.5, font.size = 10)

# Visualize dominant senders (sources) and receivers (targets) in scatterplot
# Signaling role analysis on aggregated network from all signaling pathways
setwd(wd)
tiff("CellChat_ScatterPlot_IGF_VEGF.tiff", width=500, height=500, res=120)
netAnalysis_signalingRole_scatter(cellchat, signaling = c("IGF", "VEGF")) # Signaling role analysis on networks of interest
dev.off()

# Identify signals contributing most to outgoing or incoming signaling
setwd(wd)
tiff("CellChat_Outgoing_IGF_VEGF.tiff", width=1000, height=800, res=120)
netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", signaling = c("IGF", "VEGF"))
tiff("CellChat_Incomingt_IGF_VEGF.tiff", width=1000, height=800, res=120)
netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", signaling = c("IGF", "VEGF"))
dev.off()

# Identify global communication patterns to explore how multiple cell types and signaling pathways coordinate together

# As the number of patterns increases, there might be redundant patterns, making it difficult to interpret. We chose five patterns as default. Generally, it is biologically meaningful with number of patterns > 2. In addition, selectK can infer the number of patterns, which is based on two metrics that have been implemented in the NMF R package, including Cophenetic and Silhouette. Both metrics measure the stability for a particular number of patterns based on a hierarchical clustering of the consensus matrix. For a range of the number of patterns, a suitable number of patterns is the one at which Cophenetic and Silhouette values begin to drop suddenly (kinda like elbow plot).

# Identify outgoing communication pattern of secreting cells
# To intuitively show the associations of latent patterns with cell groups and LR pairs or signaling pathways, we used river (alluvial) plots. We first normalize each row of W and each column of H to be [0,1], and then set the elements in W and H to be zero if they are less than 0.5. This threshold uncovers the most enriched cell groups and signaling pathways associated with each inferred pattern. These thresholded matrices W and H are used as inputs for creating alluvial plots.

# To directly relate cell groups with their enriched signaling pathways, we set the elements in W and H to be zero if they are less than 1/R where R is the number of latent patterns. Using a contribution score of each cell group to each signaling pathway computed by multiplying W by H, we constructed a dot plot in which the dot size is proportion to the contribution score to show association between cell group and their enriched signaling pathways. USERS can also decrease the parameter cutoff to show more enriched signaling pathways associated each cell group.

library(NMF)
library(ggalluvial)

# Here we run selectK to infer the number of patterns.
selectK(cellchat, pattern = "outgoing")
# Both Cophenetic and Silhouette values begin to drop suddenly after the number of outgoing patterns is 6.
nPatterns <- 6
setwd(wd)
tiff("CellChat_Outgoing_Patterns.tiff", width=1000, height=1000, res=120)
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns)
dev.off()

setwd(wd)
tiff("CellChat_Outgoing_River.tiff", width=2000, height=1200, res=120)
netAnalysis_river(cellchat, pattern = "outgoing") # River plot
tiff("CellChat_Outgoing_Dot.tiff", width=2000, height=600, res=120)
netAnalysis_dot(cellchat, pattern = "outgoing") # Dot plot
dev.off()

# Identify and visualize incoming communication pattern of target cells
selectK(cellchat, pattern = "incoming")
# Cophenetic values begin to drop when the number of incoming patterns is 5.
nPatterns <- 5
setwd(wd)
tiff("CellChat_Incoming_Patterns.tiff", width=1000, height=1000, res=120)
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns)
dev.off()

setwd(wd)
tiff("CellChat_Incoming_River.tiff", width=2000, height=1200, res=120)
netAnalysis_river(cellchat, pattern = "incoming") # River plot
tiff("CellChat_Incoming_Dot.tiff", width=2000, height=600, res=120)
netAnalysis_dot(cellchat, pattern = "incoming") # Dot plot
dev.off()

# Manifold and classification learning of signaling networks
# Functional similarity: High degree of functional similarity can be interpreted as the two signaling pathways or two LR pairs exhibit similar or redundant roles. The functional similarity analysis requires the same cell population composition between two datasets.
# Structural similarity: A structural similarity was used to compare their signaling network structure, without considering the similarity of senders and receivers.

# Identify signaling groups based on their functional similarity
cellchat <- computeNetSimilarity(cellchat, type = "functional")
#pip install umap-learn --ignore-installed (run in terminal)
cellchat <- netEmbedding(cellchat, type = "functional")
cellchat <- netClustering(cellchat, type = "functional")
# Visualization in 2D-space
setwd(wd)
tiff("CellChat_Functional.tiff", width=2000, height=2000, res=120)
netVisual_embedding(cellchat, type = "functional", label.size = 3.5)
dev.off()
#netVisual_embeddingZoomIn(cellchat, type = "functional", nCol = 2)

# Identify signaling groups based on structure similarity
cellchat <- computeNetSimilarity(cellchat, type = "structural")
cellchat <- netEmbedding(cellchat, type = "structural")
cellchat <- netClustering(cellchat, type = "structural")
# Visualization in 2D-space
setwd(wd)
tiff("CellChat_Structural.tiff", width=2000, height=2000, res=120)
netVisual_embedding(cellchat, type = "structural", label.size = 3.5)
dev.off()
#netVisual_embeddingZoomIn(cellchat, type = "structural", nCol = 2)

setwd(wd)
saveRDS(cellchat, file = "cellchat_ctrl_MgEndo.rds")
```

